### 2024-10-16 paper_code.Rnw
Add files via upload

\documentclass[a4paper, 11pt]{article}
\usepackage{amsmath, amssymb} % math
\usepackage[round]{natbib} % bibliography
\usepackage[dvipsnames,table]{xcolor} % colors
\usepackage[onehalfspacing]{setspace} % more space

%% margins
\usepackage{geometry}
\geometry{
  a4paper,
  total={170mm,257mm},
  left=20mm,
  right=20mm,
  top=30mm,
  bottom=30mm,
}

\newcommand\mail{yangmi@ethz.ch}
\title{\textbf{Code exploration}}
\author{Minghan Yang}
\date{\today}

\usepackage{hyperref}
\hypersetup{
  bookmarksopen=true, 
  breaklinks=true,
  pdfsubject={},
  pdfkeywords={},
  colorlinks=true,
  linkcolor=black,
  anchorcolor=black,
  citecolor=blue,
  urlcolor=blue,
}

<< "main-setup", echo = FALSE>>=
## setting knitr options (see https://yihui.org/knitr/ for all the options)
library(knitr)
opts_chunk$set(fig.height = 4,
               echo = FALSE,
               warning = FALSE,
               message = FALSE
               )
@

\begin{document}
\SweaveOpts{concordance=TRUE}
\maketitle

\begin{abstract}
Codes related to paper "Beyond p-values: A phase II dual-criterion design with statistical significance and clinical relevance".
\end{abstract}

\section{Introduction}
Proof-of-concept in Phase II trails is important in investigating the efficacy of an experimental drug. It will influence the decision of whether continuing or not continuing the development of this drug. \\
Dual-criterion design in frequentist and Bayesian applications are discussed. \\
Three generic phase II designs are reviewed:
\begin{enumerate}
\item Standard design\\
For comparative treatment and control trails, it puts forward criteria expressed as error rates: Type I error control and Power (correctly reject $H_0$ when it is false).\\
Control type I error and maximize power.\\
Type I error: $\mathbb{P}(\text{reject } H_0 | H_0 \text{ is true})=\alpha$\\
Type II error: $\mathbb{P}(\text{not reject } H_0 | H_0 \text{ is false})=\beta$.\\
Limitation: statistical significant only guarantees evidence to reject ``No effect", but is not sufficient for clinical perspective. Also, it always result in success or failure according to statistical significance.\\
Increasing the sample size increases the power for effects better than null.
\item Dual-criterion design\\
Considers both the statistical significance and the effect estimate. \\
Required inputs: type I error control (null hypothesis and type I error $\alpha$) and a decision value (DV). The DV is same as the target difference in Fisch's paper. It is the minimal effect estimate needed for trail success (if higher than this value with moderate confidence, then GO).\\
By considering both, we have both statistical significance and guarantees a sufficiently large effect estimate.\\
The dual-criterion is more demanding, the resulting power of study is less compared to standard design.\\
Power is only increased for values superior to the DV since inferior values are clinically irrelevant.
\item Precision design\\
Doesn't rely on error rates. When null hypothesis or other benchmark values cannot be determined, this can be an option. It requires sufficiently precise effect estimate.\\
Precision=$\frac{TP}{TP+FP}$. High Precision means that when the model predicts a positive outcome, it is very likely to be correct.
\end{enumerate}
Consider hazard function in survival analysis, it describes the risk of failing. We consider Hazard ratio between experimental drug and control as the outcome of interest. Hazard ratio(HR) less than 1 means the drug is better than control. We want to reduce hazard and hazard ratio.\\
\textbf{Sample size:}
\begin{equation*}
n_{min}=\frac{\sigma^2 \times z_{\alpha}^2}{(NV-DV)^2}
\end{equation*}
It is the minimum sample size that implies statistical significance if the effect estimate equals the DV. \\

<< echo = TRUE, fig.height = 4.5 >>=
plot(Sepal.Width ~ Petal.Length, data = iris)
@

Code outputs can also be used inline with \texttt{Sexpr\{..R code..\}}, for
example, 1+1=\Sexpr{1+1}

Here is a reference \citet{Cornfield1966}

%% References
\bibliographystyle{apalike}
\bibliography{bibliography}

\newpage
\section*{Computational details}
<< "sessionInfo" >>=
cat(paste(Sys.time(), Sys.timezone(), "\n"))
sessionInfo()
@

\end{document}

### 2024-10-28 paper_code.Rnw
Update paper_code.Rnw

@@
 \documentclass[a4paper, 11pt]{article}
+\usepackage[margin=0.9in]{geometry}
 \usepackage{amsmath, amssymb} % math
@@
 \usepackage[onehalfspacing]{setspace} % more space
+\usepackage{graphicx}
+\usepackage{float}
 
 %% margins
-\usepackage{geometry}
-\geometry{
-  a4paper,
-  total={170mm,257mm},
-  left=20mm,
-  right=20mm,
-  top=30mm,
-  bottom=30mm,
-}
+% \usepackage{geometry}
+% \geometry{
+%   a4paper,
+%   total={170mm,257mm},
+%   left=20mm,
+%   right=20mm,
+%   top=30mm,
+%   bottom=30mm,
+% }
 
@@
 \begin{abstract}
-Codes related to paper "Beyond p-values: A phase II dual-criterion design with statistical significance and clinical relevance".
+Codes related to paper ``Beyond p-values: A phase II dual-criterion design with statistical significance and clinical relevance''.
 \end{abstract}
@@
 \section{Introduction}
-Proof-of-concept in Phase II trails is important in investigating the efficacy of an experimental drug.
+Proof-of-concept (POC) in Phase II trails is important in investigating the efficacy of an experimental drug.
 It will influence the decision of whether continuing or not continuing the development of this drug.
@@
 The dual-criterion is more demanding, the resulting power of study is less compared to standard design.\\
-Power is only increased for values superior to the DV since inferior values are clinically irrelevant.
+Power is only increased for values superior to the DV since inferior values are clinically irrelevant.\\
+Decisions for dual-criterion design:
+<<plot, echo=FALSE >>=
+# Load necessary libraries
+library(ggplot2)
+
+# Updated data based on visual inspection of the plot
+data <- data.frame(
+  group = factor(c("(1) NO-GO", "(2) GO", "(3)Inconclusive", "(4)Inconclusive"), 
+                 levels = c("(4)Inconclusive", "(3)Inconclusive", "(2) GO","(1) NO-GO")),
+  hazard_ratio = c(0.85, 0.6, 0.8, 0.65),     # Hazard ratio estimates for each case
+  lower_ci = c(0.55, 0.25, 0.65, 0.25),         # Lower bounds of CIs
+  upper_ci = c(1.15, 0.95, 0.95, 1.05)          # Upper bounds of CIs
+)
+
+# Plot setup
+ggplot(data, aes(x = hazard_ratio, y = group)) +
+  geom_point(shape = 18, size = 3) +   # Diamond marker for hazard ratios
+  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci), height = 0.2) + # Error bars for CIs
+  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +  # Line for HR=1
+  geom_vline(xintercept = 0.7, linetype = "dashed", color = "grey50") + # Line for DV (assumed to be 0.7)
+  annotate("text", x = 0.7, y = 0, label = "DV", vjust = -1, color = "red",fontface = "bold", size = 5) + # Label for DV
+  annotate("text", x = 1.02, y = 0, label = "1", vjust = -1, color = "red",fontface = "bold", size = 5) +   # Label for HR=1
+  labs(x = "Hazard ratio", y = "") +
+  scale_x_continuous(breaks = NULL, limits = c(0.2, 1.5)) +
+  theme_minimal(base_size = 14) +
+  theme(panel.grid.minor = element_blank()) +
+  annotate("text", x = 0.4, y = 4.5, label = "Treatment better", hjust = 0, color = "grey20") +
+  annotate("text", x = 1.4, y = 4.5, label = "Treatment worse", hjust = 1, color = "grey20")
+
+ggsave("Decisions.png", width = 8, height = 6, dpi = 300)
+@
+\begin{figure}[H]
+\centering
+\includegraphics[height=6.5cm, width=10cm]{Decisions.png}
+\caption{Decisions for dual criterion design}
+\end{figure}
 \item Precision design\\
@@
 Consider hazard function in survival analysis, it describes the risk of failing.
-We consider Hazard ratio between experimental drug and control as the outcome of interest.
+We consider hazard ratio between experimental drug and control as the outcome of interest.
 Hazard ratio(HR) less than 1 means the drug is better than control.
@@
 \end{equation*}
-It is the minimum sample size that implies statistical significance if the effect estimate equals the DV. \\
+It is the minimum sample size that implies statistical significance if the effect estimate equals the DV.
+This value is calculated under the situation that both criterion are just satisfied.
+As illustrated in the below graph, when the effect estimate $\theta=$ DV, and the lower bound of the confidence interval just touches the NV so that statistical significance is reached.
+<< plot, echo=FALSE, results=hide>>=
+png("N_min.png", width = 1200, height = 800, res = 150)
 
+# Define x-axis range and labels
+x <- c(1, 5.5)
+par(mar = c(4, 4, 2, 2) + 0.1)
+# Set up an empty plot with x-axis labeled points
+plot(1:6, type = "n", xaxt = "n", ylim = c(0.2, 0.5), yaxt = "n", ylab = "Y", xlab = "Hazard Ratio", xlim=c(-2,12))
+axis(1, at = x, labels = c("NV", "DV"))
+
+# Add vertical slim strip constrained in y-axis between 0.6 and 0.63
+rect(1, 0.3, 10, 0.32, col = rgb(0.5, 0.5, 1, alpha = 0.3), border = NA)
+
+# Add vertical dashed lines at x = 1 and x = 5.5
+abline(v = x, col = "black", lty = 2)
+abline(v=10, col="black",lty=2)
+
+library(latex2exp)
+text(x=1.3, y=0.35, labels = TeX("$\\theta-z_{\\alpha} \\sqrt{\\frac{\\sigma^2}{n_{min}}}$"))
+text(x=5.7, y=0.34, labels = TeX("$\\theta$"))
+text(x=9.7, y=0.35, labels = TeX("$\\theta+z_{\\alpha} \\sqrt{\\frac{\\sigma^2}{n_{min}}}$"))
+dev.off()
+@ 
+\begin{figure}[H]
+\centering
+\includegraphics[height=7cm, width=10cm]{N_min.png}
+\caption{An illustration of sample size calculation}
+\end{figure}
+
+\noindent \textbf{Operating characteristics:}\\
+The operating characteristics are the type I error and power of the clinical trial design.
+
+\section{Example 1: A randomized PoC design with time-to-event data}
 << echo = TRUE, fig.height = 4.5 >>=
-plot(Sepal.Width ~ Petal.Length, data = iris)
+n.min <- (4*qnorm(0.9)^2)/(log(1)-log(0.7))^2
+
+# a sequence of true HR.
+t.d <- seq(0, 1, 0.1)
+
+#Dual-criterion design: alpha=0.1, DV=0.7, n=70
+n1 <- 70
+sd1 <- sqrt((2*2^2)/n1) #sigma assumed to be 2?
+cut.ssig <- 0.736 # statistical significance NV=1
+cut.crel <- 0.7 # critical relevance
+pp.go1 <- pnorm(cut.crel, t.d, sd1)
+pp.ngo1 <- 1- pnorm(cut.ssig, t.d, sd1)
+pp.intd1 <- 1 -pp.go1 - pp.ngo1
+
+#OC=65
+
+n2 <- 65
+sd2 <- sqrt((2*20^2)/n2)
+cut.ssig <- 3.179
+cut.crel <- 5
+pp.go2 <- 1- pnorm(cut.crel, t.d, sd2)
+pp.ngo2 <- pnorm(cut.ssig, t.d, sd2)
+pp.intd2 <- 1 -pp.go2 - pp.ngo2
+
+#Plot
+
+# par(mfrow=c(1,2))
+# 
+# plot(t.d, pp.go1, type="l", xlab= expression(delta), ylab="Probability", main="n=27", col="green", ylim=c(0,1))
+# lines(t.d, pp.ngo1, type="l", col="red")
+# lines(t.d, pp.intd1, type="l", col="darkgoldenrod1")
+# abline(h=0.1, lty=2)
+# abline(h=0.8, lty=2)
+# abline(v=0.8, lty=1)
+# 
+# plot(t.d, pp.go2, type="l", xlab= expression(delta), ylab="Probability", main="n=65", col="green", ylim=c(0,1))
+# lines(t.d, pp.ngo2, type="l", col="red")
+# lines(t.d, pp.intd2, type="l", col="darkgoldenrod1")
+# abline(h=0.1, lty=2)
+# abline(h=0.8, lty=2)
+# abline(v=8, lty=1)
 @
-
-Code outputs can also be used inline with \texttt{Sexpr\{..R code..\}}, for
-example, 1+1=\Sexpr{1+1}
 

### 2024-10-28 paper_code.Rnw
Update paper_code.Rnw

@@
 \begin{abstract}
-Codes related to paper ``Beyond p-values: A phase II dual-criterion design with statistical significance and clinical relevance''.
+Related to paper ``Beyond p-values: A phase II dual-criterion design with statistical significance and clinical relevance'' \citet{Roychoudhury2018}.
 \end{abstract}
@@
 As illustrated in the below graph, when the effect estimate $\theta=$ DV, and the lower bound of the confidence interval just touches the NV so that statistical significance is reached.
+Notice that when sample size equals the minimum sample size, there can only be GO or NO-GO decisions.
 << plot, echo=FALSE, results=hide>>=
@@
 # Define x-axis range and labels
-x <- c(1, 5.5)
+x <- c(5.5,10)
 par(mar = c(4, 4, 2, 2) + 0.1)
@@
 plot(1:6, type = "n", xaxt = "n", ylim = c(0.2, 0.5), yaxt = "n", ylab = "Y", xlab = "Hazard Ratio", xlim=c(-2,12))
-axis(1, at = x, labels = c("NV", "DV"))
+axis(1, at = x, labels = c("DV", "NV=1"))
 
@@
 abline(v = x, col = "black", lty = 2)
-abline(v=10, col="black",lty=2)
+abline(v=1, col="black",lty=2)
 
@@
 \noindent \textbf{Operating characteristics:}\\
-The operating characteristics are the type I error and power of the clinical trial design.
+The operating characteristics are the type I error and power of the clinical trial design. \\
+For dual-criterion designs, the power at the DV is approximately 50\%, so that if the true parameter equals the DV, there is roughly equal chance that the effect estimate lies on either side of the DV.
+Having 50\% at the DV does not mean the study is under-powered.
 
 \section{Example 1: A randomized PoC design with time-to-event data}
-<< echo = TRUE, fig.height = 4.5 >>=
+Randomized, double-blind, RCT. Patients were randomized equally to: (experimental drug + standard care) OR (standard care only).\\
+Primary outcome of interest (or called ``endpoint'') is the progression-free survival (PFS), which is the time when the disease or cancer do not get worse.
+The endpoint was assessed with a \emph{log-rank test} and \emph{Cox regression} with treatment as a covariate.
+\begin{itemize}
+\item \emph{log-rank test}: compare the survival distributions of two or more groups.
+It tests the hypothesis that there is no difference in survival (or time-to-event) between the two groups.
+If the log-rank test indicates a significant difference, it suggests the treatment affects how long patients live without their disease worsening.
+\item \emph{Cox regression}: estimate the hazard ratio between two groups, which tells us the relative risk of disease progression in the treatment group compared to the control group.
+If the $HR < 1$, it suggests that the new treatment delays disease progression better than the control treatment.
+\end{itemize}
+As for the DV, HR=0.7 was deemed necessary to be clinically meaningful. Values larger than 0.7 are unsatisfactory to clearly justify further development of the drug.\\
+So the dual-criterion is:
+\begin{enumerate}
+\item Statistical significance: one-sided p-value of log-rank test $\leq 0.1$.
+\item Clinical relevance: estimated HR from Cox regression $\leq 0.70$.
+\end{enumerate}
+\subsection{Reproduce Table 3}
+Attempt to reproduce the values in Table 3.
+The main problem is to find the correct value for $\sigma$.
+There seem to be no detailed information about the choice of value for $\sigma$.
+We chose it to be 2 according to page 455 in the paper.
+However, under this value, the calculated values do not agree with the ones in Table 3.
+\\
+Also, the calculation of $\hat{\theta} = 0.736, 0.708, 0.659, 0.761$ should be looked into when the correct $\sigma$ is decided.
+The current understanding is that these values are calculated when the estimate $\hat{\theta} > DV$ and the confidence interval just touches 1, i.e.
+NO-GO is implied.
+<< echo = FALSE >>=
 n.min <- (4*qnorm(0.9)^2)/(log(1)-log(0.7))^2
@@
 
-#Dual-criterion design: alpha=0.1, DV=0.7, n=70
+# Dual-criterion design: alpha=0.1, DV=0.7, n=70
 n1 <- 70
 sd1 <- sqrt((2*2^2)/n1) #sigma assumed to be 2?
+#sqrt((1.723522^2)/n1)
 cut.ssig <- 0.736 # statistical significance NV=1
@@
 
-#OC=65
+subtable1 <- matrix(data=c(t.d, pp.go1,pp.ngo1,pp.intd1), ncol=4)
 
-n2 <- 65
-sd2 <- sqrt((2*20^2)/n2)
-cut.ssig <- 3.179
-cut.crel <- 5
-pp.go2 <- 1- pnorm(cut.crel, t.d, sd2)
-pp.ngo2 <- pnorm(cut.ssig, t.d, sd2)
+# Dual-criterion design: alpha=0.1, DV=0.7, n=52
+n2 <- 52
+sd2 <- sqrt((2*2^2)/n2)
+cut.ssig <- 0.7
+cut.crel <- 0.7
+pp.go2 <- pnorm(cut.crel, t.d, sd2)
+pp.ngo2 <- 1-pnorm(cut.ssig, t.d, sd2)
 pp.intd2 <- 1 -pp.go2 - pp.ngo2
+
+subtable2 <- matrix(data=c(t.d, pp.go2,pp.ngo2,pp.intd2), ncol=4)
+
+# Dual-criterion design: alpha=0.1, beta=0.1, n=55
+n3 <- 55
+sd3 <- sqrt((2*2^2)/n3)
+cut.ssig <- qnorm(0.9, 0.5, sd3) # different from 0.708
+cut.crel <- qnorm(0.9, 0.5, sd3)
+pp.go3 <- pnorm(cut.crel, t.d, sd3)
+pp.ngo3 <- 1-pnorm(cut.ssig, t.d, sd3)
+pp.intd3 <- 1 -pp.go3 - pp.ngo3
+
+subtable3 <- matrix(data=c(t.d, pp.go3,pp.ngo3,pp.intd3), ncol=4)
+
+# Dual-criterion design: alpha=0.1, beta=0.2, n=38
+n4 <- 38
+sd4 <- sqrt((2*2^2)/n4)
+cut.ssig <- qnorm(0.8, 0.5, sd4) # different from 0.659
+cut.crel <- qnorm(0.8, 0.5, sd4)
+pp.go4 <- pnorm(cut.crel, t.d, sd4)
+pp.ngo4 <- 1-pnorm(cut.ssig, t.d, sd4)
+pp.intd4 <- 1 -pp.go4 - pp.ngo4
+
+subtable4 <- matrix(data=c(t.d, pp.go4,pp.ngo4,pp.intd4), ncol=4)
+
+# Dual-criterion design: alpha=0.2, beta=0.1, n=38
+n5 <- 38
+sd5 <- sqrt((2*2^2)/n5)
+cut.ssig <- qnorm(0.9, 0.5, sd5) # different from 0.659
+cut.crel <- qnorm(0.9, 0.5, sd5)
+pp.go5 <- pnorm(cut.crel, t.d, sd5)
+pp.ngo5 <- 1-pnorm(cut.ssig, t.d, sd4)
+pp.intd5 <- 1 -pp.go5 - pp.ngo5
+
+subtable5 <- matrix(data=c(t.d, pp.go5,pp.ngo5,pp.intd5), ncol=4)
+
+Table3 <- rbind(subtable1[6:11,],subtable2[6:11,],subtable3[6:11,], subtable4[6:11,], subtable5[6:11,])
+colnames(Table3) <- c("True HR", "GO", "NO-GO", "Inconclusive")
+rownames(Table3) <- c(rep("1",6),rep("2",6),rep("3",6),rep("4",6),rep("5",6))
+Table3
+@
+
+<<plots, echo = FALSE>>=
 
@@
 
-# par(mfrow=c(1,2))
-# 
-# plot(t.d, pp.go1, type="l", xlab= expression(delta), ylab="Probability", main="n=27", col="green", ylim=c(0,1))
-# lines(t.d, pp.ngo1, type="l", col="red")
-# lines(t.d, pp.intd1, type="l", col="darkgoldenrod1")
-# abline(h=0.1, lty=2)
-# abline(h=0.8, lty=2)
-# abline(v=0.8, lty=1)
-# 
-# plot(t.d, pp.go2, type="l", xlab= expression(delta), ylab="Probability", main="n=65", col="green", ylim=c(0,1))
-# lines(t.d, pp.ngo2, type="l", col="red")
-# lines(t.d, pp.intd2, type="l", col="darkgoldenrod1")
-# abline(h=0.1, lty=2)
-# abline(h=0.8, lty=2)
-# abline(v=8, lty=1)
+par(mfrow=c(1,2))
+
+plot(t.d, pp.go1, type="l", xlab= "Hazard Ratio", ylab="Probability", main="n=70", col="green", ylim=c(0,1))
+lines(t.d, pp.ngo1, type="l", col="red")
+lines(t.d, pp.intd1, type="l", col="darkgoldenrod1")
+abline(h=0.1, lty=2)
+abline(h=0.8, lty=2)
+abline(v=0.8, lty=1)
+
+plot(t.d, pp.go2, type="l", xlab= "Hazard Ratio", ylab="Probability", main="n=52", col="green", ylim=c(0,1))
+lines(t.d, pp.ngo2, type="l", col="red")
+lines(t.d, pp.intd2, type="l", col="darkgoldenrod1")
+abline(h=0.1, lty=2)
+abline(h=0.8, lty=2)
+abline(v=0.8, lty=1)
+
 @
 
-Here is a reference \citet{Cornfield1966}
+\section{Example 2: A single-arm PoC design with binary data}
+Experimental drug in Chinese patients with non-small-cell lung cancer. \\
+Primary endpoint is objective response rate (ORR), which quantifies the preliminary efficacy of the experimental drug. \\
+Prior: minimally informative unimodal beta prior distribution $Beta(0.0811,1)$, which has mean 0.75.\\
+NV is set to 7.5\% rather than 0, because of the absence of a comparator (in single arm trials). \\
+DV is set to be 10\%+7.5\%=17.5\%.\\
+So the dual-criterion is:
+\begin{enumerate}
+\item Bayesian statistical significance: $\mathbb{P}(ORR \geq 7.5\% | data) \geq 0.95$
+\item Clinical relevance: Posterior median $\geq 17.5\%$
+\end{enumerate}
+The minimal sample size was 22. Final sample size 25. \\
+Null hypothesis: there is no effect of the drug, i.e. ORR=7.5\%\\
+$\mathbb{P}(\text{type I error})=\mathbb{P}(\text{reject } H_0 | H_0 \text{ is true})=\mathbb{P}(\text{reject } H_0 | ORR \leq 7.5\%)$\\
+$\mathbb{P}(\text{type II error})=\mathbb{P}(\text{not reject } H_0 | H_0 \text{ is false})=\mathbb{P}(\text{reject } H_0 | ORR= \text{response rate})$\\
+Table 4 results show that this dual-criterion design is a three-outcome design with desirable properties.
 

### 2024-10-28 paper_code.Rnw
Update paper_code.Rnw

@@
 
+\subsection{Reproduce Figure 1}
+<<Figure 1, echo=FALSE, results=hide>>=
+# a sequence of true HR.
+t.d <- seq(0, 1, 0.01)
+
+n <- 309
+sd <- sqrt((2*2^2)/n) #sigma assumed to be 2?
+cut.ssig <- 0.8 # statistical significance NV=1
+cut.crel <- 0.8 # critical relevance
+pp.go <- pnorm(cut.crel, t.d, sd)
+pp.ngo <- 1- pnorm(cut.ssig, t.d, sd)
+pp.intd <- 1 -pp.go - pp.ngo
+
+n <- 420
+sd <- sqrt((2*2^2)/n) #sigma assumed to be 2?
+cut.ssig <- 0.83 # statistical significance NV=1
+cut.crel <- 0.8 # critical relevance
+pp.go6 <- pnorm(cut.crel, t.d, sd)
+pp.ngo6 <- 1- pnorm(cut.ssig, t.d, sd)
+pp.intd6 <- 1 -pp.go6 - pp.ngo6
+
+png("Figure 1.png", width = 1200, height = 800, res = 150)
+par(mfrow=c(1,2))
+plot(t.d, pp.go, type="l", xlab= "Hazard Ratio", ylab="Probability", main="n=309", col="green", ylim=c(0,1),xlim=c(0.3,1))
+lines(t.d, pp.ngo, type="l", col="red")
+lines(t.d, pp.intd, type="l", col="darkgoldenrod1")
+abline(v=0.8, lty=1)
+
+plot(t.d, pp.go6, type="l", xlab= "Hazard Ratio", ylab="Probability", main="n=420", col="green", ylim=c(0,1),xlim=c(0.3,1))
+lines(t.d, pp.ngo6, type="l", col="red")
+lines(t.d, pp.intd6, type="l", col="darkgoldenrod1")
+abline(v=0.8, lty=1)
+dev.off()
+
+@
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{Figure 1.png}
+\caption{Reproduce Figure 1}
+\end{figure}
+
 \section{Example 1: A randomized PoC design with time-to-event data}
@@
 
-par(mfrow=c(1,2))
-
 plot(t.d, pp.go1, type="l", xlab= "Hazard Ratio", ylab="Probability", main="n=70", col="green", ylim=c(0,1))

### 2024-10-30 paper_Roychoudhury_code.Rnw
Update paper_Roychoudhury_code.Rnw

--- 2024-10-28 paper_code.Rnw
@@
 \SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=TRUE}
 \maketitle
@@
 Decisions for dual-criterion design:
-<<plot, echo=FALSE >>=
+\begin{center}
+<<plot1, echo=FALSE, fig=TRUE>>=
 # Load necessary libraries
@@
 
-ggsave("Decisions.png", width = 8, height = 6, dpi = 300)
+#ggsave("Decisions.png", width = 8, height = 6, dpi = 300)
 @
-\begin{figure}[H]
-\centering
-\includegraphics[height=6.5cm, width=10cm]{Decisions.png}
-\caption{Decisions for dual criterion design}
-\end{figure}
+\end{center}
+%\begin{figure}[H]
+%\centering
+%\includegraphics[height=6.5cm, width=10cm]{Decisions.png}
+%\caption{Decisions for dual criterion design}
+%\end{figure}
 \item Precision design\\
 Doesn't rely on error rates. When null hypothesis or other benchmark values cannot be determined, this can be an option. It requires sufficiently precise effect estimate.\\
-Precision=$\frac{TP}{TP+FP}$. High Precision means that when the model predicts a positive outcome, it is very likely to be correct.
+Precision=$\frac{\text{TP}}{\text{TP}+\text{FP}}$. High Precision means that when the model predicts a positive outcome, it is very likely to be correct.
 \end{enumerate}
@@
 We want to reduce hazard and hazard ratio.\\
-\textbf{Sample size:}
+\textbf{Sample size:}\\
+Given the significance level $\alpha$, the null value (NV), and the decision value (DV), we can calculate the minimum sample size (for normally distributed data):
 \begin{equation*}
-n_{min}=\frac{\sigma^2 \times z_{\alpha}^2}{(NV-DV)^2}
+n_{min}=\frac{\sigma^2 \times z_{\alpha}^2}{(\text{NV}-\text{DV})^2}
 \end{equation*}
+where $\sigma$ is the outcome standard deviation, and takes the value 2 under equal randomization for the standard normal approximation to time-to-event data.
+The $z_{\alpha}$ is the $100(1-\alpha)\%$ quantile of the standard normal distribution.
 It is the minimum sample size that implies statistical significance if the effect estimate equals the DV.
 This value is calculated under the situation that both criterion are just satisfied.
-As illustrated in the below graph, when the effect estimate $\theta=$ DV, and the lower bound of the confidence interval just touches the NV so that statistical significance is reached.
+As illustrated in the below graph, when the effect estimate $\theta=$ DV, and the lower bound of the confidence interval just touches the NV so that statistical significance is reached, we have the minimum sample size.
 Notice that when sample size equals the minimum sample size, there can only be GO or NO-GO decisions.
-<< plot, echo=FALSE, results=hide>>=
-png("N_min.png", width = 1200, height = 800, res = 150)
-
+\begin{center}
+<< Minimum sample size, echo=FALSE, results=hide, fig=TRUE>>=
 # Define x-axis range and labels
@@
 library(latex2exp)
-text(x=1.3, y=0.35, labels = TeX("$\\theta-z_{\\alpha} \\sqrt{\\frac{\\sigma^2}{n_{min}}}$"))
-text(x=5.7, y=0.34, labels = TeX("$\\theta$"))
-text(x=9.7, y=0.35, labels = TeX("$\\theta+z_{\\alpha} \\sqrt{\\frac{\\sigma^2}{n_{min}}}$"))
-dev.off()
+text(x=1.3, y=0.35, labels = TeX("$\\theta-z_{\\alpha} \\sqrt{\\frac{\\sigma^2}{n_{min}}}$"),cex = 1.5)
+text(x=5.7, y=0.34, labels = TeX("$\\theta$"),cex=1.5)
+text(x=9.7, y=0.35, labels = TeX("$\\theta+z_{\\alpha} \\sqrt{\\frac{\\sigma^2}{n_{min}}}$"),cex=1.5)
+text(x=5, y=0.31, labels = "confidence interval", cex=1)
 @ 
-\begin{figure}[H]
-\centering
-\includegraphics[height=7cm, width=10cm]{N_min.png}
-\caption{An illustration of sample size calculation}
-\end{figure}
+\end{center}
 
@@
 Having 50\% at the DV does not mean the study is under-powered.
+
+\subsection{Reproduce Figure 1}
+In Figure 1, the two plots illustrate the operating characteristics of dual-criterion designs with 309 and 420 events.
+The number 309 is the minimum sample size calcualted under the example conditions $\sigma=2$, $\alpha=2.5\%$, log hazard ratios NV = $\log(1)$, DV = $\log(0.8)$.
+\begin{equation*}
+n_{min}=\frac{2^2 \times z_{0.025}^2}{(\log{1}-\log{0.8})^2} = 308.594 \approx 309
+\end{equation*}
+The probability of making a ``GO'' decision is the probability of the estimate smaller than the decision value and 
+<<Figure 1, eval=FALSE, echo=TRUE, results=hide, fig=TRUE, highlight=TRUE>>=
+# a sequence of true hazard ratios in log scale.
+t.d <- log(seq(0.5, 1, 0.01))
+
+# Left panel: sample size = 309
+n <- 309
+sd <- sqrt((2^2)/n) #sigma assumed to be 2. 
+cut.ssig <- log(0.8) # statistical significance 
+cut.crel <- log(0.8) # clinical relevance DV=0.8
+pp.go <- pnorm(cut.crel, t.d, sd)
+pp.ngo <- 1- pnorm(cut.ssig, t.d, sd)
+pp.intd <- 1 -pp.go - pp.ngo
+
+# Right panel: sample size = 420 
+n <- 420
+sd <- sqrt((2^2)/n) #sigma assumed to be 2
+#cut.ssig <- log(0.83) # statistical significance NV=1????
+cut.ssig <- log(1-qnorm(0.975)*sqrt(2^2/420))
+cut.crel <- log(0.8) # critical relevance
+pp.go6 <- pnorm(cut.crel, t.d, sd)
+pp.ngo6 <- 1- pnorm(cut.ssig, t.d, sd)
+pp.intd6 <- 1 -pp.go6 - pp.ngo6
+
+#png("Figure 1.png", width = 1200, height = 800, res = 150)
+par(mfrow=c(1,2),las=1)
+plot(exp(t.d), pp.go, type="l", xlab= "Hazard Ratio", ylab="Probability", 
+     main="n=309", col="green", ylim=c(0,1),xlim=c(0.5,1))
+lines(exp(t.d), pp.ngo, type="l", col="red")
+lines(exp(t.d), pp.intd, type="l", col="darkgoldenrod1")
+abline(v=0.8, lty=1)
+
+plot(exp(t.d), pp.go6, type="l", xlab= "Hazard Ratio", ylab="Probability", 
+     main="n=420", col="green", ylim=c(0,1),xlim=c(0.5,1))
+lines(exp(t.d), pp.ngo6, type="l", col="red")
+lines(exp(t.d), pp.intd6, type="l", col="darkgoldenrod1")
+abline(v=0.8, lty=1)
+@
+
+<<Figure 1, fig=TRUE>>=
+# Load necessary libraries
+library(ggplot2)
+library(cowplot)
+
+# Data preparation
+t.d <- log(seq(0.5, 1, 0.01))
+
+# Left panel data (n = 309)
+n1 <- 309
+sd1 <- sqrt((2^2) / n1)  # sigma assumed to be 2
+cut.ssig1 <- log(0.8)    # statistical significance 
+cut.crel1 <- log(0.8)    # clinical relevance DV=0.8
+pp.go1 <- pnorm(cut.crel1, t.d, sd1)
+pp.ngo1 <- 1 - pnorm(cut.ssig1, t.d, sd1)
+pp.intd1 <- 1 - pp.go1 - pp.ngo1
+
+# Right panel data (n = 420)
+n2 <- 420
+sd2 <- sqrt((2^2) / n2)  # sigma assumed to be 2
+cut.ssig2 <- log(1 - qnorm(0.975) * sqrt(2^2 / 420)) # statistical significance 
+cut.crel2 <- log(0.8)  # clinical relevance
+pp.go2 <- pnorm(cut.crel2, t.d, sd2)
+pp.ngo2 <- 1 - pnorm(cut.ssig2, t.d, sd2)
+pp.intd2 <- 1 - pp.go2 - pp.ngo2
+
+# Data for plotting
+df1 <- data.frame(HazardRatio = exp(t.d), GO = pp.go1, NOGO = pp.ngo1, 
+                  Inconclusive = pp.intd1)
+df2 <- data.frame(HazardRatio = exp(t.d), GO = pp.go2, NOGO = pp.ngo2, 
+                  Inconclusive = pp.intd2)
+
+# Left panel plot (n = 309)
+p1 <- ggplot(df1, aes(x = HazardRatio)) +
+  geom_line(aes(y = GO, color = "GO")) +
+  geom_line(aes(y = NOGO, color = "NOGO")) +
+  geom_line(aes(y = Inconclusive, color = "Inconclusive")) +
+  geom_vline(xintercept = c(0.8, 1.0), linetype = "dashed", color = "black") +  
+  labs(x = "Hazard Ratio", y = "Probability", title = "n = 309") +
+  scale_color_manual(values = c("GO" = "green", "NOGO" = "red", 
+                                "Inconclusive" = "darkgoldenrod1")) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+# Right panel plot (n = 420)
+p2 <- ggplot(df2, aes(x = HazardRatio)) +
+  geom_line(aes(y = GO, color = "GO")) +
+  geom_line(aes(y = NOGO, color = "NOGO")) +
+  geom_line(aes(y = Inconclusive, color = "Inconclusive")) +
+  geom_vline(xintercept = c(0.8, 1.0), linetype = "dashed", color = "black") +
+  labs(x = "Hazard Ratio", y = "Probability", title = "n = 420") +
+  scale_color_manual(values = c("GO" = "green", "NOGO" = "red", 
+                                "Inconclusive" = "darkgoldenrod1")) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+# Combine the plots with shared legend at the bottom
+combined_plot <- plot_grid(p1 + theme(legend.position = "none"),
+                           p2 + theme(legend.position = "none"),
+                           ncol = 2,
+                           align = 'hv',
+                           rel_widths = c(1.5, 1.5))  
+# Adjust widths to make the plots wider
+
+# Extract legend from one of the plots
+legend <- get_legend(p1)
+
+# Final plot with shared legend at the bottom
+final_plot <- plot_grid(combined_plot, legend, ncol = 1, rel_heights = c(1, 0.1))
+print(final_plot)
+@
 
@@
 \item \emph{Cox regression}: estimate the hazard ratio between two groups, which tells us the relative risk of disease progression in the treatment group compared to the control group.
-If the $HR < 1$, it suggests that the new treatment delays disease progression better than the control treatment.
+If the $\text{HR} < 1$, it suggests that the new treatment delays disease progression better than the control treatment.
 \end{itemize}
@@
 Also, the calculation of $\hat{\theta} = 0.736, 0.708, 0.659, 0.761$ should be looked into when the correct $\sigma$ is decided.
-The current understanding is that these values are calculated when the estimate $\hat{\theta} > DV$ and the confidence interval just touches 1, i.e.
+The current understanding is that these values are calculated when the estimate $\hat{\theta} > \text{DV}$ and the confidence interval just touches 1, i.e.
 NO-GO is implied.
-<< echo = FALSE >>=
+<< echo = TRUE >>=
 n.min <- (4*qnorm(0.9)^2)/(log(1)-log(0.7))^2
@@
 # a sequence of true HR.
-t.d <- seq(0, 1, 0.1)
+t.d <- log(seq(0.5, 1, 0.1))
 
@@
 n1 <- 70
-sd1 <- sqrt((2*2^2)/n1) #sigma assumed to be 2?
+sd1 <- sqrt((2^2)/n1) #sigma assumed to be 2?
 #sqrt((1.723522^2)/n1)
-cut.ssig <- 0.736 # statistical significance NV=1
-cut.crel <- 0.7 # critical relevance
+cut.ssig <- log(0.736) # statistical significance NV=1
+cut.crel <- log(0.7) # critical relevance
 pp.go1 <- pnorm(cut.crel, t.d, sd1)
@@
 n2 <- 52
-sd2 <- sqrt((2*2^2)/n2)
-cut.ssig <- 0.7
-cut.crel <- 0.7
+sd2 <- sqrt((2^2)/n2)
+cut.ssig <- log(0.7)
+cut.crel <- log(0.7)
 pp.go2 <- pnorm(cut.crel, t.d, sd2)
@@
 
-Table3 <- rbind(subtable1[6:11,],subtable2[6:11,],subtable3[6:11,], subtable4[6:11,], subtable5[6:11,])
-colnames(Table3) <- c("True HR", "GO", "NO-GO", "Inconclusive")
-rownames(Table3) <- c(rep("1",6),rep("2",6),rep("3",6),rep("4",6),rep("5",6))
-Table3
+# Table3 <- rbind(subtable1[6:11,],subtable2[6:11,],subtable3[6:11,], 
+# subtable4[6:11,], subtable5[6:11,])
+# colnames(Table3) <- c("True HR", "GO", "NO-GO", "Inconclusive")
+# rownames(Table3) <- c(rep("1",6),rep("2",6),rep("3",6),rep("4",6),rep("5",6))
+# Table3
 @
@@
 
-par(mfrow=c(1,2))
-
-plot(t.d, pp.go1, type="l", xlab= "Hazard Ratio", ylab="Probability", main="n=70", col="green", ylim=c(0,1))
+plot(t.d, pp.go1, type="l", xlab= "Hazard Ratio", ylab="Probability", 
+     main="n=70", col="green", ylim=c(0,1))
 lines(t.d, pp.ngo1, type="l", col="red")
@@
 
-plot(t.d, pp.go2, type="l", xlab= "Hazard Ratio", ylab="Probability", main="n=52", col="green", ylim=c(0,1))
+plot(t.d, pp.go2, type="l", xlab= "Hazard Ratio", ylab="Probability", 
+     main="n=52", col="green", ylim=c(0,1))
 lines(t.d, pp.ngo2, type="l", col="red")
@@
 
+\subsection{Reproduce Figure 2}
+<<Figure 2, echo=FALSE>>=
+
+@
+
+\subsection{Reproduce Table 4}
+<<Table 4, echo=FALSE>>=
+
+@
+
 %% References

### 2024-11-03 paper_Roychoudhury_code.Rnw
Update paper_Roychoudhury_code.Rnw

@@
 \usepackage{float}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
 
@@
 The probability of making a ``GO'' decision is the probability of the estimate smaller than the decision value and 
-<<Figure 1, eval=FALSE, echo=TRUE, results=hide, fig=TRUE, highlight=TRUE>>=
-# a sequence of true hazard ratios in log scale.
-t.d <- log(seq(0.5, 1, 0.01))
-
-# Left panel: sample size = 309
-n <- 309
-sd <- sqrt((2^2)/n) #sigma assumed to be 2. 
-cut.ssig <- log(0.8) # statistical significance 
-cut.crel <- log(0.8) # clinical relevance DV=0.8
-pp.go <- pnorm(cut.crel, t.d, sd)
-pp.ngo <- 1- pnorm(cut.ssig, t.d, sd)
-pp.intd <- 1 -pp.go - pp.ngo
-
-# Right panel: sample size = 420 
-n <- 420
-sd <- sqrt((2^2)/n) #sigma assumed to be 2
-#cut.ssig <- log(0.83) # statistical significance NV=1????
-cut.ssig <- log(1-qnorm(0.975)*sqrt(2^2/420))
-cut.crel <- log(0.8) # critical relevance
-pp.go6 <- pnorm(cut.crel, t.d, sd)
-pp.ngo6 <- 1- pnorm(cut.ssig, t.d, sd)
-pp.intd6 <- 1 -pp.go6 - pp.ngo6
-
-#png("Figure 1.png", width = 1200, height = 800, res = 150)
-par(mfrow=c(1,2),las=1)
-plot(exp(t.d), pp.go, type="l", xlab= "Hazard Ratio", ylab="Probability", 
-     main="n=309", col="green", ylim=c(0,1),xlim=c(0.5,1))
-lines(exp(t.d), pp.ngo, type="l", col="red")
-lines(exp(t.d), pp.intd, type="l", col="darkgoldenrod1")
-abline(v=0.8, lty=1)
-
-plot(exp(t.d), pp.go6, type="l", xlab= "Hazard Ratio", ylab="Probability", 
-     main="n=420", col="green", ylim=c(0,1),xlim=c(0.5,1))
-lines(exp(t.d), pp.ngo6, type="l", col="red")
-lines(exp(t.d), pp.intd6, type="l", col="darkgoldenrod1")
-abline(v=0.8, lty=1)
-@
-
-<<Figure 1, fig=TRUE>>=
+\begin{center}
+<<Figure 1, fig=TRUE, results=hide>>=
 # Load necessary libraries
@@
 library(cowplot)
+library(gridExtra)
+library(grid)
 
-# Data preparation
+# Sequence of true hazard ratios in log scale
 t.d <- log(seq(0.5, 1, 0.01))
 
-# Left panel data (n = 309)
+# Left panel (n = 309)
 n1 <- 309
-sd1 <- sqrt((2^2) / n1)  # sigma assumed to be 2
-cut.ssig1 <- log(0.8)    # statistical significance 
-cut.crel1 <- log(0.8)    # clinical relevance DV=0.8
+sd1 <- sqrt((2^2) / n1)
+cut.ssig1 <- log(0.8)
+cut.crel1 <- log(0.8)
 pp.go1 <- pnorm(cut.crel1, t.d, sd1)
@@
 pp.intd1 <- 1 - pp.go1 - pp.ngo1
+df1 <- data.frame(HazardRatio = exp(t.d), GO = pp.go1, NOGO = pp.ngo1, Inconclusive = pp.intd1)
 
-# Right panel data (n = 420)
+# Right panel (n = 420)
 n2 <- 420
-sd2 <- sqrt((2^2) / n2)  # sigma assumed to be 2
-cut.ssig2 <- log(1 - qnorm(0.975) * sqrt(2^2 / 420)) # statistical significance 
-cut.crel2 <- log(0.8)  # clinical relevance
+sd2 <- sqrt((2^2) / n2)
+cut.ssig2 <- log(1) - qnorm(0.975) * sqrt(2^2 / 420)
+cut.crel2 <- log(0.8)
 pp.go2 <- pnorm(cut.crel2, t.d, sd2)
@@
 pp.intd2 <- 1 - pp.go2 - pp.ngo2
+df2 <- data.frame(HazardRatio = exp(t.d), GO = pp.go2, NOGO = pp.ngo2, Inconclusive = pp.intd2)
 
-# Data for plotting
-df1 <- data.frame(HazardRatio = exp(t.d), GO = pp.go1, NOGO = pp.ngo1, 
-                  Inconclusive = pp.intd1)
-df2 <- data.frame(HazardRatio = exp(t.d), GO = pp.go2, NOGO = pp.ngo2, 
-                  Inconclusive = pp.intd2)
+# Define the line colors and types
+line_colors <- c("GO" = "green3", "Inconclusive" = "darkgoldenrod2", "NO-GO" = "red3")
+line_types <- c("GO" = "solid", "Inconclusive" = "dotted", "NO-GO" = "dashed")
 
-# Left panel plot (n = 309)
+# Plot for n = 309
 p1 <- ggplot(df1, aes(x = HazardRatio)) +
-  geom_line(aes(y = GO, color = "GO")) +
-  geom_line(aes(y = NOGO, color = "NOGO")) +
-  geom_line(aes(y = Inconclusive, color = "Inconclusive")) +
+  geom_line(aes(y = GO, color = "GO", linetype = "GO"),size=0.8) +
+  geom_line(aes(y = NOGO, color = "NO-GO", linetype = "NO-GO"),size=0.8) +
+  geom_line(aes(y = Inconclusive, color = "Inconclusive", 
+                linetype = "Inconclusive"),size=1.5) +
   geom_vline(xintercept = c(0.8, 1.0), linetype = "dashed", color = "black") +
-  labs(x = "Hazard Ratio", y = "Probability", title = "n = 309") +
-  scale_color_manual(values = c("GO" = "green", "NOGO" = "red", 
-                                "Inconclusive" = "darkgoldenrod1")) +
+  labs(title = "Nevent = 309") +
+  scale_color_manual(values = line_colors) +
+  scale_linetype_manual(values = line_types) +
   theme_minimal() +
-  theme(legend.position = "bottom")
+  theme(
+    legend.position = "none",
+    plot.title = element_text(hjust = 0.5, face = "bold"),
+    panel.grid.minor = element_blank(),
+    axis.title.y = element_blank(),  # Remove individual y-labels
+    axis.title.x = element_blank()
+  )
 
-# Right panel plot (n = 420)
+# Plot for n = 420
 p2 <- ggplot(df2, aes(x = HazardRatio)) +
-  geom_line(aes(y = GO, color = "GO")) +
-  geom_line(aes(y = NOGO, color = "NOGO")) +
-  geom_line(aes(y = Inconclusive, color = "Inconclusive")) +
+  geom_line(aes(y = GO, color = "GO", linetype = "GO"),size=0.8) +
+  geom_line(aes(y = NOGO, color = "NO-GO", linetype = "NO-GO"),size=0.8) +
+  geom_line(aes(y = Inconclusive, color = "Inconclusive", 
+                linetype = "Inconclusive"),size=1.5) +
   geom_vline(xintercept = c(0.8, 1.0), linetype = "dashed", color = "black") +
-  labs(x = "Hazard Ratio", y = "Probability", title = "n = 420") +
-  scale_color_manual(values = c("GO" = "green", "NOGO" = "red", 
-                                "Inconclusive" = "darkgoldenrod1")) +
+  labs(title = "Nevent = 420") +
+  scale_color_manual(values = line_colors) +
+  scale_linetype_manual(values = line_types) +
   theme_minimal() +
-  theme(legend.position = "bottom")
+  theme(
+    legend.position = "none",
+    plot.title = element_text(hjust = 0.5, face = "bold"),
+    panel.grid.minor = element_blank(),
+    axis.title.y = element_blank(),  # Remove individual y-labels
+    axis.title.x = element_blank()
+  )
 
-# Combine the plots with shared legend at the bottom
-combined_plot <- plot_grid(p1 + theme(legend.position = "none"),
-                           p2 + theme(legend.position = "none"),
-                           ncol = 2,
-                           align = 'hv',
-                           rel_widths = c(1.5, 1.5))  
-# Adjust widths to make the plots wider
+# Combine the plots into a single figure without individual y-axis labels
+combined_plots <- plot_grid(p1, p2, ncol = 2, align = 'hv', rel_widths = c(1, 1))
 
-# Extract legend from one of the plots
-legend <- get_legend(p1)
+# Extract and create a shared legend
+legend <- get_legend(
+  p1 + theme(legend.position = "right") +
+    guides(color = guide_legend(title = "Decision", nrow = 1), 
+           linetype = guide_legend(title = "Decision", nrow = 1)))
 
-# Final plot with shared legend at the bottom
-final_plot <- plot_grid(combined_plot, legend, ncol = 1, rel_heights = c(1, 0.1))
+# Add a shared y-axis label using grid.arrange
+final_plot <- grid.arrange(
+  arrangeGrob(combined_plots, 
+              left = textGrob("Probability", rot = 90, vjust = 1.2),
+              bottom = textGrob("Hazard Ratio",just = "centre")),
+  legend = legend,
+  ncol = 1,
+  heights = c(10, 1) 
+)
+
+# Print the final plot
 print(final_plot)
 @
+\end{center}
 
@@
 We chose it to be 2 according to page 455 in the paper.
-However, under this value, the calculated values do not agree with the ones in Table 3.
 \\
@@
 NO-GO is implied.
-<< echo = TRUE >>=
+<< Table 3, echo = TRUE, results=tex >>=
+library(kableExtra)
 n.min <- (4*qnorm(0.9)^2)/(log(1)-log(0.7))^2
@@
 #sqrt((1.723522^2)/n1)
-cut.ssig <- log(0.736) # statistical significance NV=1
+# cut.ssig <- 0-qnorm(0.9)* sqrt(2^2 / 70)
+cut.ssig <- log(exp(0-qnorm(0.9)* sqrt(2^2 / n1))) # statistical significance NV=1
 cut.crel <- log(0.7) # critical relevance
@@
 
-subtable1 <- matrix(data=c(t.d, pp.go1,pp.ngo1,pp.intd1), ncol=4)
+subtable1 <- matrix(data=round(c(exp(t.d), pp.go1,pp.ngo1,pp.intd1),3), ncol=4)
 
@@
 
-subtable2 <- matrix(data=c(t.d, pp.go2,pp.ngo2,pp.intd2), ncol=4)
+subtable2 <- matrix(data=round(c(exp(t.d), pp.go2,pp.ngo2,pp.intd2),3), ncol=4)
 
@@
 n3 <- 55
-sd3 <- sqrt((2*2^2)/n3)
-cut.ssig <- qnorm(0.9, 0.5, sd3) # different from 0.708
-cut.crel <- qnorm(0.9, 0.5, sd3)
+sd3 <- sqrt((2^2)/n3)
+cut.ssig <- log(0.708)
+cut.crel <- log(0.708)
+cut.ssig <- qnorm(0.9,log(0.5),sd3) # different from 0.708
+cut.crel <- qnorm(0.9,log(0.5),sd3)
 pp.go3 <- pnorm(cut.crel, t.d, sd3)
@@
 
-subtable3 <- matrix(data=c(t.d, pp.go3,pp.ngo3,pp.intd3), ncol=4)
+subtable3 <- matrix(data=round(c(exp(t.d), pp.go3,pp.ngo3,pp.intd3),3), ncol=4)
 
@@
 n4 <- 38
-sd4 <- sqrt((2*2^2)/n4)
-cut.ssig <- qnorm(0.8, 0.5, sd4) # different from 0.659
-cut.crel <- qnorm(0.8, 0.5, sd4)
+sd4 <- sqrt((2^2)/n4)
+cut.ssig <- log(exp(0-qnorm(0.9)* sqrt(2^2 / n4))) # different from 0.659
+cut.crel <- log(exp(0-qnorm(0.9)* sqrt(2^2 / n4)))
 pp.go4 <- pnorm(cut.crel, t.d, sd4)
@@
 
-subtable4 <- matrix(data=c(t.d, pp.go4,pp.ngo4,pp.intd4), ncol=4)
+subtable4 <- matrix(data=round(c(exp(t.d), pp.go4,pp.ngo4,pp.intd4),3), ncol=4)
 
@@
 n5 <- 38
-sd5 <- sqrt((2*2^2)/n5)
-cut.ssig <- qnorm(0.9, 0.5, sd5) # different from 0.659
-cut.crel <- qnorm(0.9, 0.5, sd5)
+sd5 <- sqrt((2^2)/n5)
+cut.ssig <- log(exp(0-qnorm(0.8)* sqrt(2^2 / n5))) 
+# How is the power guaranteed to be 0.1? This is only using alpha=0.2.
+cut.crel <- log(exp(0-qnorm(0.8)* sqrt(2^2 / n5)))
 pp.go5 <- pnorm(cut.crel, t.d, sd5)
-pp.ngo5 <- 1-pnorm(cut.ssig, t.d, sd4)
+pp.ngo5 <- 1-pnorm(cut.ssig, t.d, sd5)
 pp.intd5 <- 1 -pp.go5 - pp.ngo5
 
-subtable5 <- matrix(data=c(t.d, pp.go5,pp.ngo5,pp.intd5), ncol=4)
+subtable5 <- matrix(data=round(c(exp(t.d), pp.go5,pp.ngo5,pp.intd5),3), ncol=4)
 
-# Table3 <- rbind(subtable1[6:11,],subtable2[6:11,],subtable3[6:11,], 
-# subtable4[6:11,], subtable5[6:11,])
-# colnames(Table3) <- c("True HR", "GO", "NO-GO", "Inconclusive")
-# rownames(Table3) <- c(rep("1",6),rep("2",6),rep("3",6),rep("4",6),rep("5",6))
-# Table3
+# Combine subtables by rows
+combined_table <- rbind(subtable1, subtable2, subtable3, subtable4, subtable5)
+
+# Convert to data frame for better kable support
+combined_table <- as.data.frame(combined_table)
+
+kable(combined_table, format = "latex", align = "c", booktabs = TRUE,
+      col.names = c("True HR", "GO", "NO-GO", "Inconclusive")) %>%
+  kable_styling(full_width = FALSE, position = "center",
+                latex_options = c("hold_position","scale_down")) %>%
+  add_header_above(c("Reproduced Table 3" = 4)) %>%
+  group_rows("Subtable 1", 1, 6) %>%
+  group_rows("Subtable 2", 7, 12) %>%
+  group_rows("Subtable 3", 13, 18) %>%
+  group_rows("Subtable 4", 19, 24) %>%
+  group_rows("Subtable 5", 25, 30)
 @
 
-<<plots, echo = FALSE>>=
-
-#Plot
-
-plot(t.d, pp.go1, type="l", xlab= "Hazard Ratio", ylab="Probability", 
-     main="n=70", col="green", ylim=c(0,1))
-lines(t.d, pp.ngo1, type="l", col="red")
-lines(t.d, pp.intd1, type="l", col="darkgoldenrod1")
-abline(h=0.1, lty=2)
-abline(h=0.8, lty=2)
-abline(v=0.8, lty=1)
-
-plot(t.d, pp.go2, type="l", xlab= "Hazard Ratio", ylab="Probability", 
-     main="n=52", col="green", ylim=c(0,1))
-lines(t.d, pp.ngo2, type="l", col="red")
-lines(t.d, pp.intd2, type="l", col="darkgoldenrod1")
-abline(h=0.1, lty=2)
-abline(h=0.8, lty=2)
-abline(v=0.8, lty=1)
-
-@
-
+\pagebreak
 \section{Example 2: A single-arm PoC design with binary data}

### 2024-11-07 paper_Roychoudhury_code.Rnw
Update paper_Roychoudhury_code.Rnw

@@
                )
+
+# Load necessary libraries
+library(ggplot2)
+library(cowplot)
+library(gridExtra)
+library(grid)
+library(kableExtra)
 @
@@
 <<plot1, echo=FALSE, fig=TRUE>>=
-# Load necessary libraries
-library(ggplot2)
-
 # Updated data based on visual inspection of the plot
@@
   annotate("text", x = 1.4, y = 4.5, label = "Treatment worse", hjust = 1, color = "grey20")
-
-#ggsave("Decisions.png", width = 8, height = 6, dpi = 300)
 @
 \end{center}
-%\begin{figure}[H]
-%\centering
-%\includegraphics[height=6.5cm, width=10cm]{Decisions.png}
-%\caption{Decisions for dual criterion design}
-%\end{figure}
+
 \item Precision design\\
@@
 \end{enumerate}
+
+\subsection{Hazard ratio and log hazard ratio}
 Consider hazard function in survival analysis, it describes the risk of failing.
 We consider hazard ratio between experimental drug and control as the outcome of interest.
-Hazard ratio(HR) less than 1 means the drug is better than control.
-We want to reduce hazard and hazard ratio.\\
-\textbf{Sample size:}\\
+Hazard ratio(HR) less than 1 means the drug is better than the control.
+We want to have smaller hazard and hazard ratio, so that the drug is more effective.
+\\
+In the paper, the log hazard ratio ($\log$ HR) is used instead of the hazard ratio. This could be because of the following reasons.\\
+Firstly, according to the Central Limit Theorem, as the sample size increases, the distribution of the estimator (the log HR) approaches a normal distribution.
+Also, the HR itself is a ratio of hazard rates and is positively skewed, meaning it doesn’t naturally fit a normal distribution.
+Taking the logarithm of the HR stabilizes its variance, transforming the skewed HR distribution into one that is more symmetric and closer to normal.
+Moreover, this approximate normality of $\log$ HR is useful in sample size calculation, which is discussed in the next session.
+
+\subsection{Sample size}
 Given the significance level $\alpha$, the null value (NV), and the decision value (DV), we can calculate the minimum sample size (for normally distributed data):
 \begin{equation*}
-n_{min}=\frac{\sigma^2 \times z_{\alpha}^2}{(\text{NV}-\text{DV})^2}
+n_{\text{min}}=\frac{\sigma^2 \times z_{\alpha}^2}{(\text{NV}-\text{DV})^2}
 \end{equation*}
 where $\sigma$ is the outcome standard deviation, and takes the value 2 under equal randomization for the standard normal approximation to time-to-event data. 
+%In the log scale, if we say the value is within $\pm 2$ about the truth, then this translates into a range of approximately $\pm 100 \time 2 \%$ about the truth on the original scale. 
 The $z_{\alpha}$ is the $100(1-\alpha)\%$ quantile of the standard normal distribution.
-It is the minimum sample size that implies statistical significance if the effect estimate equals the DV.
+The $n_{\text{min}}$ gives the minimum sample size that implies statistical significance if the effect estimate equals the DV.
 This value is calculated under the situation that both criterion are just satisfied.
 As illustrated in the below graph, when the effect estimate $\theta=$ DV, and the lower bound of the confidence interval just touches the NV so that statistical significance is reached, we have the minimum sample size.
-Notice that when sample size equals the minimum sample size, there can only be GO or NO-GO decisions.
+Notice that when sample size equals the minimum sample size, the width of the confidence interval $z_{\alpha}\sqrt{\frac{\sigma^2}{n_{\text{min}}}}$ equals NV-DV, so there will be no ``Inconclusive'' decisions.
+When the sample size is larger, the confidence interval becomes narrower, then an ``Inconclusive'' decision will occur.
 \begin{center}
@@
 
-\noindent \textbf{Operating characteristics:}\\
+\subsection{Operating characteristics}
 The operating characteristics are the type I error and power of the clinical trial design. \\
@@
 
-\subsection{Reproduce Figure 1}
+\subsection{Reproduce Figure 1} \label{1.4}
 In Figure 1, the two plots illustrate the operating characteristics of dual-criterion designs with 309 and 420 events.
-The number 309 is the minimum sample size calcualted under the example conditions $\sigma=2$, $\alpha=2.5\%$, log hazard ratios NV = $\log(1)$, DV = $\log(0.8)$.
+The number 309 is the minimum sample size calcualted under the example conditions $\sigma=2$, $\alpha=2.5\%$, $\log$ hazard ratios NV = $\log(1)$, DV = $\log(0.8)$.
 \begin{equation*}
-n_{min}=\frac{2^2 \times z_{0.025}^2}{(\log{1}-\log{0.8})^2} = 308.594 \approx 309
+n_{\text{min}}=\frac{2^2 \times z_{0.025}^2}{(\log{1}-\log{0.8})^2} = 308.594 \approx 309
 \end{equation*}
-The probability of making a ``GO'' decision is the probability of the estimate smaller than the decision value and 
+The probability of making a ``GO'' decision is the probability of the estimate smaller than the DV (i.e.
+clinical relevance) while the NV is outside the confidence interval (i.e.
+statistical significance).
+The ``NO-GO'' decision is made when the estimate is larger than the DV, and the NV is inside the confidence interval.
+The ``Inconclusive'' decisions are made if neither ``GO'' nor ``NO-GO'' is satisfied.
+The probability of ``Inconclusive'' decision is hence 1 minus the probability of ``GO'' and ``NO-GO'' decisions.
+\\
+Below code presents the process of obtaining Figure 1 in the paper \citet{Roychoudhury2018}.
+An extra notice here is the calculation of the cutting value for statistical significance when the sample size is larger than $n_{\text{min}}$.
+As we mentioned earlier, when $n > n_{\text{min}}$, the confidence interval is shorter, so the cutting value \texttt{cut.ssig} of the ``NO-GO'' decision should be a value between DV and NV such that $z_{\alpha}\sqrt{\frac{\sigma^2}{n}} = \text{NV} - \texttt{cut.ssig}$.
+So in this case, the $\texttt{cut.ssig} = \log(1) - z_{\alpha}\sqrt{\frac{\sigma^2}{n}}$.
 \begin{center}
 <<Figure 1, fig=TRUE, results=hide>>=
-# Load necessary libraries
-library(ggplot2)
-library(cowplot)
-library(gridExtra)
-library(grid)
-
 # Sequence of true hazard ratios in log scale
@@
 n1 <- 309
-sd1 <- sqrt((2^2) / n1)
-cut.ssig1 <- log(0.8)
-cut.crel1 <- log(0.8)
-pp.go1 <- pnorm(cut.crel1, t.d, sd1)
-pp.ngo1 <- 1 - pnorm(cut.ssig1, t.d, sd1)
-pp.intd1 <- 1 - pp.go1 - pp.ngo1
-df1 <- data.frame(HazardRatio = exp(t.d), GO = pp.go1, NOGO = pp.ngo1, Inconclusive = pp.intd1)
+sd1 <- sqrt((2^2) / n1) # standard deviation
+cut.ssig1 <- log(0.8) # cutting point for statistical significance
+cut.crel1 <- log(0.8) # cutting point for clinical relevance
+pp.go1 <- pnorm(cut.crel1, t.d, sd1) # probability of GO decision
+pp.ngo1 <- 1 - pnorm(cut.ssig1, t.d, sd1) # probability of NO-GO decision
+pp.intd1 <- 1 - pp.go1 - pp.ngo1 # probability of inconclusive decision
+df1 <- data.frame(HazardRatio = exp(t.d), GO = pp.go1, 
+                  NOGO = pp.ngo1, Inconclusive = pp.intd1)
 
@@
 n2 <- 420
-sd2 <- sqrt((2^2) / n2)
-cut.ssig2 <- log(1) - qnorm(0.975) * sqrt(2^2 / 420)
-cut.crel2 <- log(0.8)
-pp.go2 <- pnorm(cut.crel2, t.d, sd2)
-pp.ngo2 <- 1 - pnorm(cut.ssig2, t.d, sd2)
-pp.intd2 <- 1 - pp.go2 - pp.ngo2
-df2 <- data.frame(HazardRatio = exp(t.d), GO = pp.go2, NOGO = pp.ngo2, Inconclusive = pp.intd2)
+sd2 <- sqrt((2^2) / n2) # standard deviation
+cut.ssig2 <- log(1) - qnorm(0.975) * sqrt(2^2 / 420) # statistical significance
+cut.crel2 <- log(0.8) # clinical relevance
+pp.go2 <- pnorm(cut.crel2, t.d, sd2) # probability of GO decision
+pp.ngo2 <- 1 - pnorm(cut.ssig2, t.d, sd2) # probability of NO-GO decision
+pp.intd2 <- 1 - pp.go2 - pp.ngo2 # probability of inconclusive decision
+df2 <- data.frame(HazardRatio = exp(t.d), GO = pp.go2, 
+                  NOGO = pp.ngo2, Inconclusive = pp.intd2)
 
@@
 \subsection{Reproduce Table 3}
-Attempt to reproduce the values in Table 3.
-The main problem is to find the correct value for $\sigma$.
-There seem to be no detailed information about the choice of value for $\sigma$.
-We chose it to be 2 according to page 455 in the paper.
+Here we attempt to reproduce the values in Table 3. \\
+For the first sub-table, similar to Section \ref{1.4}, when $n > n_{\text{min}}$, the cutting value $\hat{\theta}$ (\texttt{cut.ssig}) of the ``NO-GO'' decision is given by $\texttt{cut.ssig} = \log(1) - z_{\alpha}\sqrt{\frac{\sigma^2}{n}}$.
+These correspond to $\hat{\theta} > 0.736 \neq 0.7$ from the paper.
 \\
-Also, the calculation of $\hat{\theta} = 0.736, 0.708, 0.659, 0.761$ should be looked into when the correct $\sigma$ is decided.
-The current understanding is that these values are calculated when the estimate $\hat{\theta} > \text{DV}$ and the confidence interval just touches 1, i.e.
-NO-GO is implied.
+For the second sub-table, $n = n_{\text{min}}$.\\
+For the third sub-table, it requires one-sided type I error of 0.1 and power of 0.9 for HR=0.5.
+So the cutting value for clinical relevance is chosen to satisfy these requirements.
+One thing to be answered is why the authors used 0.901 as the power for HR=0.5, rather than 0.9.\\
+For the forth sub-table, it requires one-sided type I error of 0.1 and power of 0.8 for HR=0.5.
+The sample size $n=38 < n_{\text{min}}$.
+The cutting value of the decisions are calculated in a similar way as in Section \ref{1.4}.
+One thing to be answered here is that the current calculation returns the same values in the table, but the $\hat{\theta}$ is slightly different from 0.659 as claimed in the paper.
+Also, the power at HR=0.5 is 0.804, rather than 0.8.
+\\
+For sub-table five, it used a type I error of 0.2 and a power of 0.9, with sample size $n=38 < n_{\text{min}}$.
+The cutting value of the decisions are calculated in a similar way as in Section \ref{1.4}.
+However the question to be answered is the same with that in sub-table 4.
 << Table 3, echo = TRUE, results=tex >>=
-library(kableExtra)
-n.min <- (4*qnorm(0.9)^2)/(log(1)-log(0.7))^2
+# minimum sample size
+n.min <- ceiling((4*qnorm(0.9)^2)/(log(1)-log(0.7))^2) # n.min = 52
 
-# a sequence of true HR.
+# a sequence of true log(HR).
 t.d <- log(seq(0.5, 1, 0.1))
 
-# Dual-criterion design: alpha=0.1, DV=0.7, n=70
+# Dual-criterion design: alpha=0.1, DV=log(0.7), n=70
 n1 <- 70
-sd1 <- sqrt((2^2)/n1) #sigma assumed to be 2?
-#sqrt((1.723522^2)/n1)
-# cut.ssig <- 0-qnorm(0.9)* sqrt(2^2 / 70)
-cut.ssig <- log(exp(0-qnorm(0.9)* sqrt(2^2 / n1))) # statistical significance NV=1
+sd1 <- sqrt((2^2)/n1) # standard deviation
+cut.ssig <- log(1)-qnorm(0.9)* sqrt(2^2 / n1) # statistical significance
 cut.crel <- log(0.7) # critical relevance
@@
 sd3 <- sqrt((2^2)/n3)
-cut.ssig <- log(0.708)
-cut.crel <- log(0.708)
-cut.ssig <- qnorm(0.9,log(0.5),sd3) # different from 0.708
-cut.crel <- qnorm(0.9,log(0.5),sd3)
+cut.ssig <- qnorm(0.901,log(0.5),sd3) # reason for using 0.901 rather than 0.9?
+cut.crel <- qnorm(0.901,log(0.5),sd3)
 pp.go3 <- pnorm(cut.crel, t.d, sd3)
@@
 sd4 <- sqrt((2^2)/n4)
-cut.ssig <- log(exp(0-qnorm(0.9)* sqrt(2^2 / n4))) # different from 0.659
-cut.crel <- log(exp(0-qnorm(0.9)* sqrt(2^2 / n4)))
+cut.ssig <- log(1)-qnorm(0.9)* sqrt(2^2 / n4) # different from 0.659
+cut.crel <- log(1)-qnorm(0.9)* sqrt(2^2 / n4)
 pp.go4 <- pnorm(cut.crel, t.d, sd4)
@@
 sd5 <- sqrt((2^2)/n5)
-cut.ssig <- log(exp(0-qnorm(0.8)* sqrt(2^2 / n5))) 
+cut.ssig <- log(1)-qnorm(0.8)* sqrt(2^2 / n5)
 # How is the power guaranteed to be 0.1? This is only using alpha=0.2.
-cut.crel <- log(exp(0-qnorm(0.8)* sqrt(2^2 / n5)))
+cut.crel <- log(1)-qnorm(0.8)* sqrt(2^2 / n5)
 pp.go5 <- pnorm(cut.crel, t.d, sd5)

### 2024-11-10 paper_Roychoudhury_code.Rnw
Update paper_Roychoudhury_code.Rnw

@@
 \usepackage{array}
+\usepackage[labelformat = empty]{caption}
 
@@
 library(knitr)
+opts_knit$set(concordance = FALSE)
 opts_chunk$set(fig.height = 4,
@@
 library(kableExtra)
+library(latex2exp)
 @
@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=FALSE}
 %\SweaveOpts{concordance=TRUE}
@@
 \begin{center}
-<<plot1, echo=FALSE, fig=TRUE>>=
+<<plot1, echo=FALSE>>=
 # Updated data based on visual inspection of the plot
@@
   group = factor(c("(1) NO-GO", "(2) GO", "(3)Inconclusive", "(4)Inconclusive"), 
-                 levels = c("(4)Inconclusive", "(3)Inconclusive", "(2) GO","(1) NO-GO")),
+                 levels = c("(4)Inconclusive", "(3)Inconclusive", 
+                            "(2) GO","(1) NO-GO")),
   hazard_ratio = c(0.85, 0.6, 0.8, 0.65),     # Hazard ratio estimates for each case
@@
 \begin{center}
-<< Minimum sample size, echo=FALSE, results=hide, fig=TRUE>>=
+<< Minimum_sample_size, echo=FALSE>>=
 # Define x-axis range and labels
@@
 # Set up an empty plot with x-axis labeled points
-plot(1:6, type = "n", xaxt = "n", ylim = c(0.2, 0.5), yaxt = "n", ylab = "Y", xlab = "Hazard Ratio", xlim=c(-2,12))
+plot(1:6, type = "n", xaxt = "n", ylim = c(0.2, 0.5), yaxt = "n", ylab = "", xlab = "Hazard Ratio", xlim=c(-2,12))
 axis(1, at = x, labels = c("DV", "NV=1"))
@@
 
-library(latex2exp)
 text(x=1.3, y=0.35, labels = TeX("$\\theta-z_{\\alpha} \\sqrt{\\frac{\\sigma^2}{n_{min}}}$"),cex = 1.5)
@@
 \begin{center}
-<<Figure 1, fig=TRUE, results=hide>>=
+<<Figure_1, results='hide', fig.height=5>>=
 # Sequence of true hazard ratios in log scale
@@
 However the question to be answered is the same with that in sub-table 4.
-<< Table 3, echo = TRUE, results=tex >>=
+<< Table 3, echo = TRUE, results='asis'>>=
 # minimum sample size
@@
 subtable1 <- matrix(data=round(c(exp(t.d), pp.go1,pp.ngo1,pp.intd1),3), ncol=4)
+colnames(subtable1) <- c("True HR", "GO: $\\hat{\\theta} \\leq 0.7$",
+                         "NO-GO: $\\hat{\\theta} > 0.736$","Inconclusive")
 
@@
 subtable2 <- matrix(data=round(c(exp(t.d), pp.go2,pp.ngo2,pp.intd2),3), ncol=4)
+colnames(subtable2) <- c("True HR", "GO: $\\hat{\\theta} \\leq 0.7$",
+                         "NO-GO: $\\hat{\\theta} > 0.7$","Inconclusive")
 
@@
 subtable3 <- matrix(data=round(c(exp(t.d), pp.go3,pp.ngo3,pp.intd3),3), ncol=4)
+colnames(subtable3) <- c("True HR", "GO: $\\hat{\\theta} \\leq 0.708$",
+                         "NO-GO: $\\hat{\\theta} > 0.708$","Inconclusive")
 
@@
 subtable4 <- matrix(data=round(c(exp(t.d), pp.go4,pp.ngo4,pp.intd4),3), ncol=4)
+colnames(subtable4) <- c("True HR", "GO: $\\hat{\\theta} \\leq 0.659$",
+                         "NO-GO: $\\hat{\\theta} > 0.659$","Inconclusive")
 
@@
 subtable5 <- matrix(data=round(c(exp(t.d), pp.go5,pp.ngo5,pp.intd5),3), ncol=4)
-
-# Combine subtables by rows
-combined_table <- rbind(subtable1, subtable2, subtable3, subtable4, subtable5)
-
-# Convert to data frame for better kable support
-combined_table <- as.data.frame(combined_table)
-
-kable(combined_table, format = "latex", align = "c", booktabs = TRUE,
-      col.names = c("True HR", "GO", "NO-GO", "Inconclusive")) %>%
-  kable_styling(full_width = FALSE, position = "center",
-                latex_options = c("hold_position","scale_down")) %>%
-  add_header_above(c("Reproduced Table 3" = 4)) %>%
-  group_rows("Subtable 1", 1, 6) %>%
-  group_rows("Subtable 2", 7, 12) %>%
-  group_rows("Subtable 3", 13, 18) %>%
-  group_rows("Subtable 4", 19, 24) %>%
-  group_rows("Subtable 5", 25, 30)
+colnames(subtable5) <- c("True HR", "GO: $\\hat{\\theta} \\leq 0.761$",
+                         "NO-GO: $\\hat{\\theta} > 0.761$","Inconclusive")
 @
 
+<<Table_3, echo=FALSE, results='asis'>>=
+kable(subtable1, format = "latex", booktabs = TRUE, escape = FALSE,
+      caption = "1. Dual-criterion design: $\\alpha=0.1$, DV=0.7, n=70") %>%
+  kable_styling(latex_options = c("hold_position", "striped")) %>%
+  cat()
+kable(subtable2, format = "latex", booktabs = TRUE, escape = FALSE,
+      caption = "2. Dual-criterion design: $\\alpha=0.1$, DV=0.7, n=52") %>%
+  kable_styling(latex_options = c("hold_position", "striped")) %>%
+  cat()
+kable(subtable3, format = "latex", booktabs = TRUE, escape = FALSE,
+      caption = "3. Dual-criterion design: $\\alpha=0.1, \\beta=0.1 (\\theta_A=0.5)$, n=55") %>%
+  kable_styling(latex_options = c("hold_position", "striped")) %>%
+  cat()
+kable(subtable4, format = "latex", booktabs = TRUE, escape = FALSE,
+      caption = "4. Dual-criterion design: $\\alpha=0.1,\\beta=0.2 (\\theta_A=0.5)$, n=38") %>%
+  kable_styling(latex_options = c("hold_position", "striped")) %>%
+  cat()
+kable(subtable5, format = "latex", booktabs = TRUE, escape = FALSE,
+      caption = "5. Dual-criterion design: $\\alpha=0.2,\\beta=0.1 (\\theta_A=0.5)$, n=38") %>%
+  kable_styling(latex_options = c("hold_position", "striped")) %>%
+  cat()
+@
 \pagebreak
@@
 Primary endpoint is objective response rate (ORR), which quantifies the preliminary efficacy of the experimental drug. \\
-Prior: minimally informative unimodal beta prior distribution $Beta(0.0811,1)$, which has mean 0.75.\\
-NV is set to 7.5\% rather than 0, because of the absence of a comparator (in single arm trials). \\
-DV is set to be 10\%+7.5\%=17.5\%.\\
+Prior: minimally informative unimodal beta prior distribution $\text{Beta}(0.0811,1)$, which has mean 0.75.\\
+Because of the absence of a comparator (in single arm trials), NV is set to 7.5\% rather than 0, based on a literature review and clinical discussions.\\
+A minimum improvement of 10\% is considered necessary for further development, so DV is set to be 10\%+7.5\%=17.5\%.\\
 So the dual-criterion is:
 \begin{enumerate}
-\item Bayesian statistical significance: $\mathbb{P}(ORR \geq 7.5\% | data) \geq 0.95$
+\item Bayesian statistical significance: $\mathbb{P}(\text{ORR} \geq 7.5\% | \text{data}) \geq 0.95$
 \item Clinical relevance: Posterior median $\geq 17.5\%$
 \end{enumerate}
+
+\subsection{Reproduce Figure 2}
+For this dual-criterion, the minimally required sample size is 22. For $n \geq 22$, clinical relevance ensures statistical significance. \\
+For a Beta(a,b) prior, if we observe a number of successes (r) and failures (nn - r)
+\begin{center}
+<<Figure_2, echo=FALSE, results='hide', fig.width=5>>=
+#This function calculates the minimal sample size for Bayesian DC design with binary endpoint
+
+SS_DC_BayesBin <- function(
+  null.value, # null.value
+  p.positive  = 0.95, # required probability to be better than the null
+  dec.value,  # decision value
+  post.est = c("median","mean")[1], # median by default
+  a=1, b=1,    # prior specification, a Beta(a,b) prior
+  n.max = 1000 # numbers of iterations
+)
+{
+  NV = null.value
+  DV = dec.value
+
+  n <- 1:n.max
+
+  # required number of responders to satisfy DV criterion (posterior mean)
+  if (post.est == "mean") { # if estimate posterior mean
+    rr = rep(NA,max(n)) # a vector to store rr (required responders)
+      for ( nn in n) {
+      r = 0:nn
+      est1 = (a+r)/(a+b+nn)
+      rr[nn] = r[est1>=DV][1]
+      }
+    print(cbind(rr,n)) # required responders for each sample size n 
+    est = (a+rr)/(a+b+n) # what does it do?
+  }
+
+  if (post.est == "median") {
+    rr = rep(NA,max(n))
+    for ( nn in n) {
+      r = 0:nn
+      est1= qbeta(0.5,a+r,b+nn-r) # the median of beta distribution
+      rr[nn] = r[est1>=DV][1]
+    }
+    print(cbind(rr,n))
+    est = qbeta(0.5,a+rr,b+n-rr)
+  }
+
+  p.obs = rr/n
+
+  # find n such that posterior prob >= p.positive 
+  post.crit <- 1-pbeta(NV,a+rr,b+n-rr)
+  post.crit.index = post.crit >= p.positive
+
+  # find n such that for all m>=n, statistical criterion is fulfilled
+  find.n =  cbind( n, rev(post.crit.index), 
+                   cumsum(rev(post.crit.index))==n, rev(n))
+  colnames(find.n) = c("index","okay","all.next.okay","n")
+
+  n.min=NULL
+
+  if (any(find.n[,3]==1) ) {
+    n.okay = find.n[find.n[,3]==1,,drop=FALSE]
+    n.min = n.okay[nrow(n.okay),4]
+    n.min.10 = n.min+10
+    est=round(est,3)
+    ss.table = data.frame( n=n[1:n.min.10], # sample size
+                           r=rr[1:n.min.10], # number of responders
+                           p.obs=p.obs[1:n.min.10], #
+                           est = est[1:n.min.10], # posterior estimate of median (by default)
+                           post.p.positive=post.crit[1:n.min.10], # posterior probability of a positive effect
+                           okay=post.crit.index[1:n.min.10])
+  }
+
+  if (is.null(n.min))
+    stop("No sample size satisfies required posterior 
+         probability criterion pr(p>NV): try higher n.max")
+
+plot( 1:(n.min+10), post.crit[1:(n.min+10)], type="p", xaxs="i", xlab="n",
+        ylab="pr(p>NV)",xlim = c(0,n.min+10), ylim=c(0.5,1),
+        lwd=1,
+        #col = ifelse(post.crit.index,"green","red"),
+        pch = ifelse(post.crit.index,16,1),
+        main="prob(p>NV) as a function of n", cex.lab=1.4, cex.main=1.4)
+
+  abline(h=p.positive, lty=2)
+  abline(v=n.min, lty=3)
+  text(1.05*n.min, 0.6, paste("n.min=",n.min, sep=''),adj=0)
+
+  return(list(ss.table=ss.table,n.min=n.min))
+  }
+
+#Example
+nB <-  SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, 
+                      dec.value = 0.175, a = 0.0811, b=1,
+                      post.est = "median")
+# returns a table of minimal sample size 
+# print(nB)
+@
+\end{center}
+
 The minimal sample size was 22. Final sample size 25. \\
 Null hypothesis: there is no effect of the drug, i.e. ORR=7.5\%\\
-$\mathbb{P}(\text{type I error})=\mathbb{P}(\text{reject } H_0 | H_0 \text{ is true})=\mathbb{P}(\text{reject } H_0 | ORR \leq 7.5\%)$\\
-$\mathbb{P}(\text{type II error})=\mathbb{P}(\text{not reject } H_0 | H_0 \text{ is false})=\mathbb{P}(\text{reject } H_0 | ORR= \text{response rate})$\\
+$\mathbb{P}(\text{type I error})=\mathbb{P}(\text{reject } H_0 | H_0 \text{ is true})=\mathbb{P}(\text{reject } H_0 | \text{ORR} \leq 7.5\%)$\\
+$\mathbb{P}(\text{type II error})=\mathbb{P}(\text{not reject } H_0 | H_0 \text{ is false})=\mathbb{P}(\text{reject } H_0 | \text{ORR} = \text{response rate})$\\
 Table 4 results show that this dual-criterion design is a three-outcome design with desirable properties.
 
-\subsection{Reproduce Figure 2}
-<<Figure 2, echo=FALSE>>=
-
-@
-
 \subsection{Reproduce Table 4}
-<<Table 4, echo=FALSE>>=
-
+<<Table 4, echo=FALSE, eval=FALSE>>=
+# a vector of true ORR values
+ORR.t <- c(0.075, 0.125, 0.175, 0.225, 0.275)
 @

### 2024-11-23 paper_Roychoudhury_code.Rnw
Update paper_Roychoudhury_code.Rnw

@@
 \begin{document}
-%\SweaveOpts{concordance=TRUE}
-%\SweaveOpts{concordance=FALSE}
-%\SweaveOpts{concordance=TRUE}
-%\SweaveOpts{concordance=TRUE}
 \maketitle
@@
   annotate("text", x = 0.7, y = 0, label = "DV", vjust = -1, color = "red",fontface = "bold", size = 5) + # Label for DV
-  annotate("text", x = 1.02, y = 0, label = "1", vjust = -1, color = "red",fontface = "bold", size = 5) +   # Label for HR=1
+  annotate("text", x = 1.02, y = 0, label = "NV", vjust = -1, color = "red",fontface = "bold", size = 5) +   # Label for HR=1
   labs(x = "Hazard ratio", y = "") +
@@
 Doesn't rely on error rates. When null hypothesis or other benchmark values cannot be determined, this can be an option. It requires sufficiently precise effect estimate.\\
-Precision=$\frac{\text{TP}}{\text{TP}+\text{FP}}$. High Precision means that when the model predicts a positive outcome, it is very likely to be correct.
 \end{enumerate}
@@
 \\
-In the paper, the log hazard ratio ($\log$ HR) is used instead of the hazard ratio. This could be because of the following reasons.\\
+In the paper, the log hazard ratio $\theta$ ($\log$ HR) is used instead of the hazard ratio. This could be because of the following reasons.\\
 Firstly, according to the Central Limit Theorem, as the sample size increases, the distribution of the estimator (the log HR) approaches a normal distribution.
@@
 \subsection{Sample size}
-Given the significance level $\alpha$, the null value (NV), and the decision value (DV), we can calculate the minimum sample size (for normally distributed data):
+Given the significance level $\alpha$, the null value (NV) and the decision value (DV) for $\log$ HR, we can calculate the minimum sample size (for normally distributed data).
+Notice here that the DV $<$ NV, $\log$ DV $<$ $\log$ NV.
 \begin{equation*}
@@
 This value is calculated under the situation that both criterion are just satisfied.
-As illustrated in the below graph, when the effect estimate $\theta=$ DV, and the lower bound of the confidence interval just touches the NV so that statistical significance is reached, we have the minimum sample size.
-Notice that when sample size equals the minimum sample size, the width of the confidence interval $z_{\alpha}\sqrt{\frac{\sigma^2}{n_{\text{min}}}}$ equals NV-DV, so there will be no ``Inconclusive'' decisions.
+As illustrated in the below graph, when the effect estimate $\theta= \log$ DV, and the lower bound of the confidence interval just touches the NV so that statistical significance is reached, we have the minimum sample size.
+Notice that when sample size equals the minimum sample size, the half-width of the confidence interval $z_{\alpha}\sqrt{\frac{\sigma^2}{n_{\text{min}}}}$ equals NV-DV, so there will be no ``Inconclusive'' decisions.
 When the sample size is larger, the confidence interval becomes narrower, then an ``Inconclusive'' decision will occur.
@@
 # Set up an empty plot with x-axis labeled points
-plot(1:6, type = "n", xaxt = "n", ylim = c(0.2, 0.5), yaxt = "n", ylab = "", xlab = "Hazard Ratio", xlim=c(-2,12))
-axis(1, at = x, labels = c("DV", "NV=1"))
+plot(1:6, type = "n", xaxt = "n", ylim = c(0.2, 0.5), yaxt = "n", ylab = "", 
+     xlab = "Log Hazard Ratio", xlim=c(-2,12))
+axis(1, at = x, labels = c("DV", "NV"))
 
@@
 In Figure 1, the two plots illustrate the operating characteristics of dual-criterion designs with 309 and 420 events.
-The number 309 is the minimum sample size calcualted under the example conditions $\sigma=2$, $\alpha=2.5\%$, $\log$ hazard ratios NV = $\log(1)$, DV = $\log(0.8)$.
+The number 309 is the minimum sample size calculated under the example conditions $\sigma=2$, $\alpha=2.5\%$, NV = $\log(1)$ and DV = $\log(0.8)$ for $\log$ HR.
 \begin{equation*}
@@
 Below code presents the process of obtaining Figure 1 in the paper \citet{Roychoudhury2018}.
-An extra notice here is the calculation of the cutting value for statistical significance when the sample size is larger than $n_{\text{min}}$.
-As we mentioned earlier, when $n > n_{\text{min}}$, the confidence interval is shorter, so the cutting value \texttt{cut.ssig} of the ``NO-GO'' decision should be a value between DV and NV such that $z_{\alpha}\sqrt{\frac{\sigma^2}{n}} = \text{NV} - \texttt{cut.ssig}$.
+An extra notice here is the calculation of the cut value for statistical significance when the sample size is larger than $n_{\text{min}}$.
+As we mentioned earlier, when $n > n_{\text{min}}$, the confidence interval is shorter, so the cut value \texttt{cut.ssig} of the ``NO-GO'' decision should be a value between DV and NV such that $z_{\alpha}\sqrt{\frac{\sigma^2}{n}} = \text{NV} - \texttt{cut.ssig}$.
 So in this case, the $\texttt{cut.ssig} = \log(1) - z_{\alpha}\sqrt{\frac{\sigma^2}{n}}$.
 \begin{center}
-<<Figure_1, results='hide', fig.height=5>>=
+<<Figure_1, results='hide', fig.height=5, echo=TRUE>>=
 # Sequence of true hazard ratios in log scale
@@
 Here we attempt to reproduce the values in Table 3. \\
-For the first sub-table, similar to Section \ref{1.4}, when $n > n_{\text{min}}$, the cutting value $\hat{\theta}$ (\texttt{cut.ssig}) of the ``NO-GO'' decision is given by $\texttt{cut.ssig} = \log(1) - z_{\alpha}\sqrt{\frac{\sigma^2}{n}}$.
+For the first sub-table, similar to Section \ref{1.4}, when $n > n_{\text{min}}$, the cut value $\hat{\theta}$ (\texttt{cut.ssig}) of the ``NO-GO'' decision is given by $\texttt{cut.ssig} = \log(1) - z_{\alpha}\sqrt{\frac{\sigma^2}{n}}$.
 These correspond to $\hat{\theta} > 0.736 \neq 0.7$ from the paper.
@@
 For the third sub-table, it requires one-sided type I error of 0.1 and power of 0.9 for HR=0.5.
-So the cutting value for clinical relevance is chosen to satisfy these requirements.
-One thing to be answered is why the authors used 0.901 as the power for HR=0.5, rather than 0.9.\\
+So the cut value for clinical relevance is chosen to satisfy these requirements.
+The authors used 0.901 as the power for HR=0.5, rather than 0.9, probably because this ensures a power of at least 0.9.\\
 For the forth sub-table, it requires one-sided type I error of 0.1 and power of 0.8 for HR=0.5.
 The sample size $n=38 < n_{\text{min}}$.
-The cutting value of the decisions are calculated in a similar way as in Section \ref{1.4}.
-One thing to be answered here is that the current calculation returns the same values in the table, but the $\hat{\theta}$ is slightly different from 0.659 as claimed in the paper.
-Also, the power at HR=0.5 is 0.804, rather than 0.8.
-\\
-For sub-table five, it used a type I error of 0.2 and a power of 0.9, with sample size $n=38 < n_{\text{min}}$.
-The cutting value of the decisions are calculated in a similar way as in Section \ref{1.4}.
-However the question to be answered is the same with that in sub-table 4.
-<< Table 3, echo = TRUE, results='asis'>>=
+The cut value of the decisions are calculated in a similar way as in Section \ref{1.4}.
+The cut value of $\hat{\theta}$ is slightly different from 0.659 as claimed in the paper, very possibly because of rounding.\\
+For sub-table five, it used a type I error of 0.2 and a power of 0.9, with sample size $n=38 < n_{\text{min}}$. The cut value of the decisions are calculated in a similar way as in Section \ref{1.4}.
+<< Table_3_code, echo = TRUE, results='asis'>>=
 # minimum sample size
@@
 sd3 <- sqrt((2^2)/n3)
-cut.ssig <- qnorm(0.901,log(0.5),sd3) # reason for using 0.901 rather than 0.9?
+cut.ssig <- qnorm(0.901,log(0.5),sd3) 
 cut.crel <- qnorm(0.901,log(0.5),sd3)
@@
 cut.ssig <- log(1)-qnorm(0.8)* sqrt(2^2 / n5)
-# How is the power guaranteed to be 0.1? This is only using alpha=0.2.
 cut.crel <- log(1)-qnorm(0.8)* sqrt(2^2 / n5)
@@
 Primary endpoint is objective response rate (ORR), which quantifies the preliminary efficacy of the experimental drug. \\
-Prior: minimally informative unimodal beta prior distribution $\text{Beta}(0.0811,1)$, which has mean 0.75.\\
+Prior: minimally informative unimodal beta prior distribution $\text{Beta}(0.0811,1)$, which has mean 0.075.\\
 Because of the absence of a comparator (in single arm trials), NV is set to 7.5\% rather than 0, based on a literature review and clinical discussions.\\
-A minimum improvement of 10\% is considered necessary for further development, so DV is set to be 10\%+7.5\%=17.5\%.\\
+A minimum improvement of 10\% is considered necessary for further development, so DV is set to be 10\%+7.5\%=17.5\%. Notice in this example, the NV < DV.\\
 So the dual-criterion is:
@@
 \end{enumerate}
+Null hypothesis: there is no effect of the drug, i.e. ORR=7.5\%\\
+$\mathbb{P}(\text{type I error})=\mathbb{P}(\text{reject } H_0 | H_0 \text{ is true})=\mathbb{P}(\text{reject } H_0 | \text{ORR} \leq 7.5\%)$\\
+$\mathbb{P}(\text{type II error})=\mathbb{P}(\text{not reject } H_0 | H_0 \text{ is false})=\mathbb{P}(\text{reject } H_0 | \text{ORR} = \text{response rate})$
 
 \subsection{Reproduce Figure 2}
-For this dual-criterion, the minimally required sample size is 22. For $n \geq 22$, clinical relevance ensures statistical significance. \\
-For a Beta(a,b) prior, if we observe a number of successes (r) and failures (nn - r)
+For this dual-criterion, the minimally required sample size is 22.
+For $n \geq 22$, clinical relevance ensures statistical significance.
+To see this, we are now going to reproduce Figure 2 in the paper.\\
+For a Beta($a$,$b$) prior, 
+\begin{equation*}
+\text{Beta}(a,b) = \frac{\theta^{a-1}(1-\theta)^{b-1}}{B(a,b)}
+\end{equation*}
+with binomial sampling (according to the paper), if we observe a number of successes ($r$) and failures ($n - r$), then the posterior is Beta($a + r$, $b + n - r$).\\
+In the below code, we first find the required numbers of responders (\texttt{rr}) to satisfy clinical relevance under different sample sizes ($n$), then we find the minimal sample size $n$ such that clinical relevance guarantees Bayesian statistical significance.
+\\
+The minimal sample size is 22. Final sample size is taken to be 25. \\
 \begin{center}
-<<Figure_2, echo=FALSE, results='hide', fig.width=5>>=
-#This function calculates the minimal sample size for Bayesian DC design with binary endpoint
-
+<<Figure_2, results='hide', fig.width=5, echo=TRUE>>=
+#This function calculates the minimal sample size 
+# for Bayesian DC design with binary endpoint
 SS_DC_BayesBin <- function(
@@
   n.max = 1000 # numbers of iterations
-)
-{
+){
   NV = null.value
   DV = dec.value
-
   n <- 1:n.max
@@
     rr = rep(NA,max(n)) # a vector to store rr (required responders)
-      for ( nn in n) {
-      r = 0:nn
-      est1 = (a+r)/(a+b+nn)
-      rr[nn] = r[est1>=DV][1]
+      for ( nn in n) { # n iterations
+      r = 0:nn # different numbers of responders
+      est1 = (a+r)/(a+b+nn) # posterior estimate based on different success numbers
+      rr[nn] = r[est1>=DV][1] # minimum r s.t. DV criterion is fulfilled
       }
     print(cbind(rr,n)) # required responders for each sample size n 
-    est = (a+rr)/(a+b+n) # what does it do?
+    est = (a+rr)/(a+b+n) # posterior estimate of mean
   }
 
+  # required number of responders to satisfy DV criterion (posterior median)
   if (post.est == "median") {
@@
       r = 0:nn
-      est1= qbeta(0.5,a+r,b+nn-r) # the median of beta distribution
-      rr[nn] = r[est1>=DV][1]
+      est1= qbeta(0.5,a+r,b+nn-r) # the median of beta posterior
+      rr[nn] = r[est1>=DV][1] # minimum r s.t. posterior median >= DV
     }
@@
 
-  # find n such that posterior prob >= p.positive 
-  post.crit <- 1-pbeta(NV,a+rr,b+n-rr)
-  post.crit.index = post.crit >= p.positive
+  # find n such that posterior prob >= p.positive (= 0.95)
+  # in other words, find n such that statistical significance is guaranteed.
+  post.crit <- 1 - pbeta(NV,a+rr,b+n-rr) # posterior probability of being greater than NV.
+  post.crit.index = post.crit >= p.positive # index=TRUE if the probability >= 0.95
 
@@
   find.n =  cbind( n, rev(post.crit.index), 
-                   cumsum(rev(post.crit.index))==n, rev(n))
+                   cumsum(rev(post.crit.index))==n, 
+                   # true if all m>=n fulfill stats criterion
+                   rev(n))
   colnames(find.n) = c("index","okay","all.next.okay","n")
@@
   if (any(find.n[,3]==1) ) {
-    n.okay = find.n[find.n[,3]==1,,drop=FALSE]
-    n.min = n.okay[nrow(n.okay),4]
-    n.min.10 = n.min+10
+    n.okay = find.n[find.n[,3]==1,,drop=FALSE] # take all the rows with all.next.okay==1
+    n.min = n.okay[nrow(n.okay),4] # take the n in the last row (find minimum n)
+    n.min.10 = n.min+15
     est=round(est,3)
-    ss.table = data.frame( n=n[1:n.min.10], # sample size
-                           r=rr[1:n.min.10], # number of responders
-                           p.obs=p.obs[1:n.min.10], #
-                           est = est[1:n.min.10], # posterior estimate of median (by default)
-                           post.p.positive=post.crit[1:n.min.10], # posterior probability of a positive effect
-                           okay=post.crit.index[1:n.min.10])
-  }
+    ss.table = data.frame( n=n[1:n.min.10], 
+                           # sample size from 1 to n.min+10
+                           r=rr[1:n.min.10], 
+                           # number of responders needed for DV criterion
+                           p.obs=p.obs[1:n.min.10], 
+                           est = est[1:n.min.10], 
+                           # posterior estimate of median (by default)
+                           post.p.positive=post.crit[1:n.min.10], 
+                           # posterior probability of a positive effect
+                           okay=post.crit.index[1:n.min.10]
+                           # TRUE if statistical significance is satisfied
+                           )}
 
@@
 
-plot( 1:(n.min+10), post.crit[1:(n.min+10)], type="p", xaxs="i", xlab="n",
-        ylab="pr(p>NV)",xlim = c(0,n.min+10), ylim=c(0.5,1),
-        lwd=1,
-        #col = ifelse(post.crit.index,"green","red"),
-        pch = ifelse(post.crit.index,16,1),
+plot( 1:(n.min+15), post.crit[1:(n.min+15)], type="p", xaxs="i", xlab="n",
+        ylab="pr(p>NV)",xlim = c(0,n.min+20), ylim=c(0.5,1),
+        lwd=1, pch = ifelse(post.crit.index,16,1),
         main="prob(p>NV) as a function of n", cex.lab=1.4, cex.main=1.4)
-
   abline(h=p.positive, lty=2)
@@
 
-  return(list(ss.table=ss.table,n.min=n.min))
-  }
+  return(list(ss.table=ss.table,n.min=n.min))}
 
@@
                       post.est = "median")
+@
+\end{center}
+\begin{table}[H]
+\centering
+<<Table_Figure_2, echo=TRUE, results='asis'>>=
 # returns a table of minimal sample size 
-# print(nB)
+kable(nB$ss.table)
+@
+\caption*{Additional Table: Minimal sample size required for the analysis.}
+\label{Additional Table}
+\end{table}
+\pagebreak
+Now we explore how the prior choice affect the minimum sample size.
+Firstly, it might be more reasonable to choose the prior with median 0.075, rather than with mean 0.075.
+So we try Beta(0.2676,1) as the prior, as it has median 0.075.
+The below codes and plot explore how this affects the minimum sample size.
+From the plot we can spot some slight difference compared to the previous plot, but their difference is not significant.
+\begin{center}
+<<prior_sample_size1, fig.width=5, echo=TRUE, results='hide'>>=
+
+# First try beta prior with median 0.075, rather than mean 0.075
+# Target median
+median_target <- 0.075
+beta_param <- 1
+
+# Function to find alpha such that the median of Beta(alpha, beta=1) is 0.075
+find_alpha <- function(alpha) {
+  abs(pbeta(median_target, alpha, beta_param) - 0.5)
+}
+
+# Use optimization to find alpha
+alpha_result <- optimize(find_alpha, c(0.001, 10))$minimum
+alpha_result # 0.2676044
+
+nM <-  SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, 
+                      dec.value = 0.175, a = alpha_result, b=1,
+                      post.est = "median")
+nM$n.min # is still 22
+@
+\end{center}
+Additionally, we explore how the minimum sample size changes according to different values of $\alpha$. The below plot gives the relationship between them.
+\begin{center}
+<<prior_sample_size2, fig.show='asis', echo=FALSE>>=
+# Next plot the minimum sample size VS alpha
+
+alpha_vector <- seq(0.005,1,0.005)
+# n_min_vector <- rep(0, length(alpha_vector))
+# for (i in 1:length(alpha_vector)){
+#   a <- alpha_vector[i]
+#   n <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, 
+#                       dec.value = 0.175, a = a, b=1,
+#                       post.est = "median")
+#   n_min_vector[i] <- n$n.min[[1]]
+# }
+# n_min_vector
+
+n_min_vector_stored <- c(21,21,21,21,21,21,21,21,21,21,21,21,21,21,21,
+                         22,22,22,22,22,22,22,22,22,22,22,22,22,22,22, 
+                         22,22,22,22,22,22,22,22,22,22,22,22,22,22,22,
+                         22,22,22,22,22,22,22,22,22,17,17,17,23,23,23,
+                         23,23,23,23,23,23,23,23,23,23,23,23,23,23,23,
+                         23,23,23,23,23,23,23,17,17,17,17,17,18,18,18,
+                         18,18,18,18,18,18,18,18,18,18,24,24,24,24,24,
+                         24,24,24,24,24,18,18,18,18,18,18,18,18,18,18,
+                         18,18,18,18,18,18,18,18,18,18,19,19,19,19,19,
+                         19,19,19,19,19,19,19,19,19,19,19,19,19,19,19, 
+                         19,19,19,19,19,19,19,19,19,19,19,19,19,19,19,
+                         19,19,19,19,19,19,19,20,20,20,20,20,20,20,20,
+                         20,20,20,20,20,20,20,20,20,20,20,20,20,20,20,
+                         20,20,20,20,20)
+plot(alpha_vector,n_min_vector_stored, type="l",
+     xlab="alpha", ylab="n_min", las=1)
 @
@@
 
-The minimal sample size was 22. Final sample size 25. \\
-Null hypothesis: there is no effect of the drug, i.e. ORR=7.5\%\\
-$\mathbb{P}(\text{type I error})=\mathbb{P}(\text{reject } H_0 | H_0 \text{ is true})=\mathbb{P}(\text{reject } H_0 | \text{ORR} \leq 7.5\%)$\\
-$\mathbb{P}(\text{type II error})=\mathbb{P}(\text{not reject } H_0 | H_0 \text{ is false})=\mathbb{P}(\text{reject } H_0 | \text{ORR} = \text{response rate})$\\
-Table 4 results show that this dual-criterion design is a three-outcome design with desirable properties.
-
 \subsection{Reproduce Table 4}
-<<Table 4, echo=FALSE, eval=FALSE>>=
+Table 4 results show that this dual-criterion design is a three-outcome design with desirable properties. \\
+Introduction of the three-outcome design \citet{SARGENT}.\\
+A design with three possible outcomes includes an additional region, called the inconclusive region, between these two regions.
+This requires two cut-off points, $r$ and $s$, with the following decision rule\\
+If $Y \leq r$, reject $H_1$\\
+If $r < Y < s$, inconclusive\\
+If $s \leq Y$, reject $H_0$\\
+Define the following parameters
+\begin{align*}
+\alpha &= \max_{p \in (0,p_0]} \mathbb{P}(\text{reject } H_0 | H_0 \text{ true}) = \mathbb{P}(Y \geq s | p = p_0)\\
+\beta &= \max_{p \in [p_1,1)} \mathbb{P}(\text{reject } H_1 | H_1 \text{ true}) = \mathbb{P}(Y \leq r | p = p_1)\\
+\eta &= \max_{p \in (0,p_0]} \mathbb{P}(\text{reject } H_1 | H_0 \text{ true}) = \mathbb{P}(Y \leq r | p = p_0)\\
+\pi &= \max_{p \in [p_1,1)} \mathbb{P}(\text{reject } H_0 | H_1 \text{ true}) = \mathbb{P}(Y \geq s | p = p_1)
+\end{align*}
+where $\alpha$ is the maximum probability of making a wrong decision by rejecting the null hypothesis when it is true, and $\eta$ is the minimum probability of making the decision to reject alternative hypothesis when the null is true.
+In contrast to the standard two-outcome design, $\eta + \alpha <1$ due to the presence of the inconclusive region.
+The same interpretation holds true for $\pi$ and $\beta$ relative to the alternative hypothesis.\\
+Define $\lambda := \mathbb{P}(r < Y < s | p = p_0)$ and $\delta := \mathbb{P}(r < Y < s | p = p_1)$, then we have 
+\begin{equation*}
+\eta + \lambda + \alpha = 1, \quad \quad
+\beta + \delta + \pi = 1
+\end{equation*}
+In the paper ``Beyond p-values: A phase II dual-criterion design with statistical significance and clinical relevance'', they consider a three-outcome design with $H_0: \text{ORR} \leq 7.5\%$, $H_1: \text{ORR} \geq 27.5\%$, $\alpha = 0.05$, $\beta = 0.1$, $\eta = 0.8$, $\pi = 0.9$, and the minimal sample size is 27.
+\\
+The below code reproduce Table 4.
+The first design is the situation when sample size is the minimum sample size, at which the inconclusive case does not happen.
+The second design has larger sample size ($n=36$), where an inconclusive case will happen.
+For a ``GO'' decision, we need the number of responders to be larger or equal to 7 according to \hyperref[Additional Table]{Additional Table} in the sample size discussion, so that the clinical relevance guarantees the statistical significance.
+For ``NO-GO'' decision, the maximum number of responders are chosen when the statistical significance is just not satisfied.
+The third design is the three-outcome design, where the cut values of numbers of responders are chosen based on the values of $\eta$ and $\alpha$.
+<<Table_4_code, echo=TRUE, results='asis'>>=
 # a vector of true ORR values
 ORR.t <- c(0.075, 0.125, 0.175, 0.225, 0.275)
+a = 0.0811
+b=1
+
+# Dual-criterion design: alpha=0.05, DV=0.175, n=25
+n1 <- 25 # is the minimum sample size
+DV <- 0.175
+NV <- 0.075
+alpha <- 0.05
+
+# required number of responders for statistical stignificance
+cut.ssig <- nB$ss.table$r[nB$ss.table$n==n1] 
+# required number of responders for clinical relevance
+cut.crel <- nB$ss.table$r[nB$ss.table$n==n1]
+
+# calculate probabilities
+p_GO1 <- round(1-pbinom(cut.crel-1, n1, prob=ORR.t, lower.tail = T),3)
+p_NOGO1 <- round(pbinom(cut.ssig-1, n1, prob=ORR.t, lower.tail = T),3)
+p_inconcl1 <- round(1- p_GO1 - p_NOGO1,3)
+
+subtable_1 <- matrix(data=c(ORR.t*100, round(c(p_GO1,p_NOGO1,p_inconcl1),3)), ncol=4)
+colnames(subtable_1) <- c("True ORR(\\%)", "GO: $r \\geq 5$",
+                         "NO-GO: $r < 5$","Inconclusive")
+
+# Dual-criterion design: alpha=0.05, DV=0.175, n=36
+n2 <- 36
+DV <- 0.175
+NV <- 0.075
+alpha <- 0.05
+
+# required number of responders for clinical relevance
+cut.crel <- nB$ss.table$r[nB$ss.table$n==n2] 
+# find cut.ssig: required number of responders for statistical significance
+r <- seq(0,10,1)
+crit <- 1 - pbeta(NV,a+r,b+n2-r) # posterior probability of being less than NV.
+crit.index = crit < 1-alpha
+cut.ssig <- max(r[crit.index])
+
+# calculate probabilities
+p_GO2 <- round(1-pbinom(cut.crel-1, n2, prob=ORR.t, lower.tail = T),3)
+p_NOGO2 <- round(pbinom(cut.ssig, n2, prob=ORR.t, lower.tail = T),3)
+p_inconcl2 <- round(1- p_GO2 - p_NOGO2,3)
+
+subtable_2 <- matrix(data=c(ORR.t*100, round(c(p_GO2,p_NOGO2,p_inconcl2),3)), ncol=4)
+colnames(subtable_2) <- c("True ORR(\\%)", "GO: $r \\geq 7$",
+                         "NO-GO: $r \\leq 5$","Inconclusive: r = 6")
+
+# Three-outsome design: n=27, H0:ORR<0.075, H1:ORR>=0.275, 
+# alpha=0.05, beta=0.1, eta=0.8, pi=0.9
+n3 <- 27 # is the minimum sample size under three-criterion design
+eta <- 0.8
+alpha <- 0.05
+
+# find s and r for three-outcome design
+s <- qbinom(alpha, n3, prob=0.075, lower.tail = F)+1
+r <- qbinom(eta, n3, prob=0.075, lower.tail = T)
+
+# calculate probabilities
+p_GO3 <- round(1-pbinom(s-1, n3, prob=ORR.t, lower.tail = T),3) # Y >= s
+p_NOGO3 <- round(pbinom(r, n3, prob=ORR.t, lower.tail = T),3) # Y <= r
+p_inconcl3 <- round(1- p_GO3 - p_NOGO3, 3)
+
+subtable_3 <- matrix(data=c(ORR.t*100, round(c(p_GO3,p_NOGO3,p_inconcl3),3)), ncol=4)
+colnames(subtable_3) <- c("True ORR(\\%)", "GO: $r \\geq 5$",
+                         "NO-GO: $r \\leq 3$","Inconclusive: r = 4")
+
 @
 
+<<Table_4, echo=FALSE, results='asis'>>=
+kable(subtable_1, format = "latex", booktabs = TRUE, escape = FALSE,
+      caption = "1. Dual-criterion design: $\\alpha=0.1$, DV=0.7, n=70") %>%
+  kable_styling(latex_options = c("hold_position", "striped")) %>%
+  cat()
+kable(subtable_2, format = "latex", booktabs = TRUE, escape = FALSE,
+      caption = "2. Dual-criterion design: $\\alpha=0.1$, DV=0.7, n=52") %>%
+  kable_styling(latex_options = c("hold_position", "striped")) %>%
+  cat()
+kable(subtable_3, format = "latex", booktabs = TRUE, escape = FALSE,
+      caption = "3. Three-outcome design: $n=27$") %>%
+  kable_styling(latex_options = c("hold_position", "striped")) %>%
+  cat()
+@
+\pagebreak
 %% References

### 2024-12-04 paper_Roychoudhury_code.Rnw
Update paper_Roychoudhury_code.Rnw

@@
   text(1.05*n.min, 0.6, paste("n.min=",n.min, sep=''),adj=0)
+  text(30, 0.8, paste("a=",round(a,4)))
 
@@
 nM$n.min # is still 22
+
 @
@@
 
-alpha_vector <- seq(0.005,1,0.005)
+# alpha_vector <- seq(0.1,1,0.1)
 # n_min_vector <- rep(0, length(alpha_vector))
+# par(mfrow=c(1,1))
 # for (i in 1:length(alpha_vector)){
@@
 # n_min_vector
-
+# 
+# 
+alpha_vector <- seq(0.005,1,0.005)
 n_min_vector_stored <- c(21,21,21,21,21,21,21,21,21,21,21,21,21,21,21,
@@
 %% References
+\appendix
+
+<<appendix1, fig.width=5, echo=TRUE, results='hide'>>=
+n1 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = 0.1, b=1, post.est = "median")
+n2 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = 0.2, b=1, post.est = "median")
+n3 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = 0.3, b=1, post.est = "median")
+n4 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = 0.4, b=1, post.est = "median")
+n5 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = 0.5, b=1, post.est = "median")
+n6 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = 0.6, b=1, post.est = "median")
+n7 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = 0.7, b=1, post.est = "median")
+n8 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = 0.8, b=1, post.est = "median")
+n9 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = 0.9, b=1, post.est = "median")
+@
 \bibliographystyle{apalike}

### 2024-12-18 paper_Roychoudhury_code.Rnw
Update paper_Roychoudhury_code.Rnw

--- 2024-11-23 paper_Roychoudhury_code.Rnw
@@
     n.min = n.okay[nrow(n.okay),4] # take the n in the last row (find minimum n)
-    n.min.10 = n.min+15
+    n.min.15 = n.min+15
     est=round(est,3)
-    ss.table = data.frame( n=n[1:n.min.10], 
+    ss.table = data.frame( n=n[1:n.min.15], 
                            # sample size from 1 to n.min+10
-                           r=rr[1:n.min.10], 
+                           r=rr[1:n.min.15], 
                            # number of responders needed for DV criterion
-                           p.obs=p.obs[1:n.min.10], 
-                           est = est[1:n.min.10], 
+                           p.obs=p.obs[1:n.min.15], 
+                           est = est[1:n.min.15], 
                            # posterior estimate of median (by default)
-                           post.p.positive=post.crit[1:n.min.10], 
+                           post.p.positive=post.crit[1:n.min.15], 
                            # posterior probability of a positive effect
-                           okay=post.crit.index[1:n.min.10]
+                           okay=post.crit.index[1:n.min.15]
                            # TRUE if statistical significance is satisfied
@@
 nM$n.min # is still 22
+
 @
 \end{center}
-Additionally, we explore how the minimum sample size changes according to different values of $\alpha$. The below plot gives the relationship between them.
+Additionally, we explore how the minimum sample size changes according to different values of $a$ in the Beta($a,b$) prior.
+The below plots give the relationship between the minimal sample size and the choices of $a$, with the median and mean being the posterior estimate, respectively.
+We can see from the plots that there are jumps of minimal sample sizes around some values of $a$, especially in the first plot.
+When the posterior estimate is the mean (the second plot), the jumping pattern only seem to occur for $a$ values closer to 1.
+This might suggest that the posterior mean gives more stability and less sensitivity of minimal sample size with respect to prior choices.
+For both cases, we obtain the values of $a$ at these jumping points and included relevant plots in the Appendix.
+<<prior_sample_size2, echo=TRUE, cache=TRUE, results='hide', fig.show='hide'>>=
+# Next plot the minimum sample size VS a (in Beta prior)
+alpha_vector <- seq(0.005,1,0.005)
+
+# posterior median
+n_min_vector_median <- rep(0, length(alpha_vector))
+for (i in 1:length(alpha_vector)){
+  a <- alpha_vector[i]
+  n <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95,
+                      dec.value = 0.175, a = a, b=1,
+                      post.est = "median")
+  n_min_vector_median[i] <- n$n.min[[1]]
+}
+
+# posterior mean
+n_min_vector_mean <- rep(0, length(alpha_vector))
+for (i in 1:length(alpha_vector)){
+  a <- alpha_vector[i]
+  n <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95,
+                      dec.value = 0.175, a = a, b=1,
+                      post.est = "mean")
+  n_min_vector_mean[i] <- n$n.min[[1]]
+}
+@
 \begin{center}
-<<prior_sample_size2, fig.show='asis', echo=FALSE>>=
-# Next plot the minimum sample size VS alpha
-
-alpha_vector <- seq(0.005,1,0.005)
-# n_min_vector <- rep(0, length(alpha_vector))
-# for (i in 1:length(alpha_vector)){
-#   a <- alpha_vector[i]
-#   n <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, 
-#                       dec.value = 0.175, a = a, b=1,
-#                       post.est = "median")
-#   n_min_vector[i] <- n$n.min[[1]]
-# }
-# n_min_vector
-
-n_min_vector_stored <- c(21,21,21,21,21,21,21,21,21,21,21,21,21,21,21,
-                         22,22,22,22,22,22,22,22,22,22,22,22,22,22,22, 
-                         22,22,22,22,22,22,22,22,22,22,22,22,22,22,22,
-                         22,22,22,22,22,22,22,22,22,17,17,17,23,23,23,
-                         23,23,23,23,23,23,23,23,23,23,23,23,23,23,23,
-                         23,23,23,23,23,23,23,17,17,17,17,17,18,18,18,
-                         18,18,18,18,18,18,18,18,18,18,24,24,24,24,24,
-                         24,24,24,24,24,18,18,18,18,18,18,18,18,18,18,
-                         18,18,18,18,18,18,18,18,18,18,19,19,19,19,19,
-                         19,19,19,19,19,19,19,19,19,19,19,19,19,19,19, 
-                         19,19,19,19,19,19,19,19,19,19,19,19,19,19,19,
-                         19,19,19,19,19,19,19,20,20,20,20,20,20,20,20,
-                         20,20,20,20,20,20,20,20,20,20,20,20,20,20,20,
-                         20,20,20,20,20)
-plot(alpha_vector,n_min_vector_stored, type="l",
-     xlab="alpha", ylab="n_min", las=1)
+<<prior_sample_size3, fig.show='asis'>>=
+plot(alpha_vector,n_min_vector_median, type="l",
+     xlab="alpha", ylab="n_min", las=1, main = "Posterior median")
+plot(alpha_vector,n_min_vector_mean, type="l",
+     xlab="alpha", ylab="n_min", las=1, main = "Posterior mean")
+# alpha_vector <- seq(0.005,1,0.005)
+# n_min_vector_stored <- c(21,21,21,21,21,21,21,21,21,21,21,21,21,21,21, 1
+#                          22,22,22,22,22,22,22,22,22,22,22,22,22,22,22, 2
+#                          22,22,22,22,22,22,22,22,22,22,22,22,22,22,22, 3
+#                          22,22,22,22,22,22,22,22,22,17,17,17,23,23,23, 4
+#                          23,23,23,23,23,23,23,23,23,23,23,23,23,23,23, 5
+#                          23,23,23,23,23,23,23,17,17,17,17,17,18,18,18, 6
+#                          18,18,18,18,18,18,18,18,18,18,24,24,24,24,24, 7
+#                          24,24,24,24,24,18,18,18,18,18,18,18,18,18,18, 8
+#                          18,18,18,18,18,18,18,18,18,18,19,19,19,19,19, 9
+#                          19,19,19,19,19,19,19,19,19,19,19,19,19,19,19, 10
+#                          19,19,19,19,19,19,19,19,19,19,19,19,19,19,19, 11
+#                          19,19,19,19,19,19,19,20,20,20,20,20,20,20,20, 12
+#                          20,20,20,20,20,20,20,20,20,20,20,20,20,20,20, 13
+#                          20,20,20,20,20)
 @
@@
 @
+
 \pagebreak
 %% References
+\appendix
+\section*{Appendix}
+\begin{center}
+Posterior median
+<<appendix1, fig.width=10, echo=TRUE, results='hide'>>=
+# First get the values of a in Beta prior where the jumps in minimal sample size occurs.
+(a_vector <- alpha_vector[c(54,55,57,58,82,83,87,88, 100, 101, 110, 111, 172,173)])
+
+par(mfrow=c(1,2))
+n1 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[1], b=1, post.est = "median")
+n2 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[2], b=1, post.est = "median")
+
+n3 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[3], b=1, post.est = "median")
+n4 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[4], b=1, post.est = "median")
+
+n5 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[5], b=1, post.est = "median")
+n6 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[6], b=1, post.est = "median")
+
+n7 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[7], b=1, post.est = "median")
+n8 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[8], b=1, post.est = "median")
+
+n9 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[9], b=1, post.est = "median")
+n10 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[10], b=1, post.est = "median")
+
+n11 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[11], b=1, post.est = "median")
+n12 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[12], b=1, post.est = "median")
+
+n13 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[13], b=1, post.est = "median")
+n14 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector[14], b=1, post.est = "median")
+@
+\end{center}
+
+\begin{center}
+Posterior mean
+<<appendix2, fig.width=10, echo=TRUE, results='hide'>>=
+par(mfrow=c(1,2))
+a_vector_mean <- alpha_vector[c(164,165,175,176,191,192)]
+n1 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector_mean[1], b=1, post.est = "mean")
+n2 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector_mean[2], b=1, post.est = "mean")
+
+n3 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector_mean[3], b=1, post.est = "mean")
+n4 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector_mean[4], b=1, post.est = "mean")
+
+n5 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector_mean[5], b=1, post.est = "mean")
+n6 <- SS_DC_BayesBin(null.value = 0.075, p.positive = 0.95, dec.value = 0.175, 
+               a = a_vector_mean[6], b=1, post.est = "mean")
+@
+\end{center}
+<<appendix3, echo=FALSE, results='asis', eval=FALSE>>=
+kable(n6$ss.table, caption = paste("a=0.6, n min =",n6$n.min))
+kable(n8$ss.table, caption = paste("a=0.8, n min =",n8$n.min))
+kable(n9$ss.table, caption = paste("a=0.9, n min =",n9$n.min))
+@
+
 \bibliographystyle{apalike}

### 2025-11-21 report_Bayesian_DC.Rnw
Rename report.Rnw to report_Bayesian_DC.Rnw

--- 2024-12-18 paper_Roychoudhury_code.Rnw
@@
 \newcommand\mail{yangmi@ethz.ch}
-\title{\textbf{Code exploration}}
+\title{\textbf{Reproducible study and further explorations on Bayesian dual-criterion design}}
 \author{Minghan Yang}
@@
 
-<< "main-setup", echo = FALSE>>=
+<< "main-setup", echo = FALSE, message=FALSE, warning=FALSE>>=
 ## setting knitr options (see https://yihui.org/knitr/ for all the options)
@@
 Proof-of-concept (POC) in Phase II trails is important in investigating the efficacy of an experimental drug.
-It will influence the decision of whether continuing or not continuing the development of this drug.
+It will influence the decision of whether continuing or not on the development of the drug.
 \\
@@
 \item Standard design\\
-For comparative treatment and control trails, it puts forward criteria expressed as error rates: Type I error control and Power (correctly reject $H_0$ when it is false).\\
-Control type I error and maximize power.\\
+For comparative treatment and control trails, it puts forward criteria expressed as error rates: Type I error control and power (correctly reject $H_0$ when it is false).\\
+We want to control type I error and maximize power.\\
 Type I error: $\mathbb{P}(\text{reject } H_0 | H_0 \text{ is true})=\alpha$\\
@@
 Required inputs: type I error control (null hypothesis and type I error $\alpha$) and a decision value (DV).
-The DV is same as the target difference in Fisch's paper.
-It is the minimal effect estimate needed for trail success (if higher than this value with moderate confidence, then GO).\\
-By considering both, we have both statistical significance and guarantees a sufficiently large effect estimate.\\
-The dual-criterion is more demanding, the resulting power of study is less compared to standard design.\\
+The DV is same as the ``target difference'' in Fisch's paper.
+It is the minimal effect estimate needed for trail success (if higher than this value with moderate confidence, then a ``GO'' decision is made).\\
+By considering both criterion, we have both statistical significance and guarantees a sufficiently large effect estimate.\\
+The dual-criterion is more demanding, the resulting power of study is less than that of a standard design.\\
 Power is only increased for values superior to the DV since inferior values are clinically irrelevant.\\
@@
 Given the significance level $\alpha$, the null value (NV) and the decision value (DV) for $\log$ HR, we can calculate the minimum sample size (for normally distributed data).
-Notice here that the DV $<$ NV, $\log$ DV $<$ $\log$ NV.
+Notice here that for $\log$ HR, DV $<$ NV, where DV, NV are log hazard ratios.
 \begin{equation*}
@@
 %In the log scale, if we say the value is within $\pm 2$ about the truth, then this translates into a range of approximately $\pm 100 \time 2 \%$ about the truth on the original scale. 
-The $z_{\alpha}$ is the $100(1-\alpha)\%$ quantile of the standard normal distribution.
-The $n_{\text{min}}$ gives the minimum sample size that implies statistical significance if the effect estimate equals the DV.
+$z_{\alpha}$ is the $100(1-\alpha)\%$ quantile of the standard normal distribution.
+$n_{\text{min}}$ gives the minimum sample size that implies statistical significance if the effect estimate equals the DV.
 This value is calculated under the situation that both criterion are just satisfied.
-As illustrated in the below graph, when the effect estimate $\theta= \log$ DV, and the lower bound of the confidence interval just touches the NV so that statistical significance is reached, we have the minimum sample size.
+As illustrated in the below graph, when the effect estimate $\theta=$ DV, and the lower bound of the confidence interval just touches the NV so that statistical significance is reached, we have the minimum sample size.
 Notice that when sample size equals the minimum sample size, the half-width of the confidence interval $z_{\alpha}\sqrt{\frac{\sigma^2}{n_{\text{min}}}}$ equals NV-DV, so there will be no ``Inconclusive'' decisions.
@@
 \subsection{Operating characteristics}
-The operating characteristics are the type I error and power of the clinical trial design. \\
+The operating characteristics are the type I error and the power of the clinical trial design. \\
 For dual-criterion designs, the power at the DV is approximately 50\%, so that if the true parameter equals the DV, there is roughly equal chance that the effect estimate lies on either side of the DV.
@@
 In Figure 1, the two plots illustrate the operating characteristics of dual-criterion designs with 309 and 420 events.
-The number 309 is the minimum sample size calculated under the example conditions $\sigma=2$, $\alpha=2.5\%$, NV = $\log(1)$ and DV = $\log(0.8)$ for $\log$ HR.
+The number 309 is the minimum sample size calculated under the example conditions $\sigma=2$, $\alpha=2.5\%$, NV = $\log(1)$ and DV = $\log(0.8)$ for $\log$ HR:
 \begin{equation*}
-n_{\text{min}}=\frac{2^2 \times z_{0.025}^2}{(\log{1}-\log{0.8})^2} = 308.594 \approx 309
+n_{\text{min}}=\frac{2^2 \times z_{0.025}^2}{(\log{1}-\log{0.8})^2} = 308.594 \approx 309.
 \end{equation*}
@@
 Primary endpoint is objective response rate (ORR), which quantifies the preliminary efficacy of the experimental drug. \\
-Prior: minimally informative unimodal beta prior distribution $\text{Beta}(0.0811,1)$, which has mean 0.075.\\
-Because of the absence of a comparator (in single arm trials), NV is set to 7.5\% rather than 0, based on a literature review and clinical discussions.\\
-A minimum improvement of 10\% is considered necessary for further development, so DV is set to be 10\%+7.5\%=17.5\%. Notice in this example, the NV < DV.\\
+Because of the absence of a comparator (in single arm trials), NV is set to 7.5\% rather than 0, based on a literature review and clinical discussions.
+Hence the prior is chosen as a minimally informative unimodal beta prior distribution $\text{Beta}(0.0811,1)$, which gives a mean of 0.075.\\
+A minimum improvement of 10\% is considered necessary for further development, so the DV is set to be 10\%+7.5\%=17.5\%. Notice in this example, NV $<$ DV.\\
 So the dual-criterion is:

### 2025-11-21 report_Group_Seq.Rnw
Rename Exploration_on_Gsponer_paper.Rnw to report_Group_Seq.Rnw

\documentclass[a4paper, 11pt]{article}
\usepackage[margin=0.9in]{geometry}
\usepackage{amsmath, amssymb} % math
\usepackage[round]{natbib} % bibliography
\usepackage[dvipsnames,table]{xcolor} % colors
\usepackage[onehalfspacing]{setspace} % more space
\usepackage{graphicx}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage[labelformat = empty]{caption}
\usepackage{xcolor}
\include{newCommands}

%% margins
% \usepackage{geometry}
% \geometry{
%   a4paper,
%   total={170mm,257mm},
%   left=20mm,
%   right=20mm,
%   top=30mm,
%   bottom=30mm,
% }

\newcommand\mail{yangmi@ethz.ch}
\title{\textbf{Exploration on Bayesian Group Sequential Designs}}
\author{Minghan Yang}
\date{April 2025}

\usepackage{hyperref}
\hypersetup{
  bookmarksopen=true, 
  breaklinks=true,
  pdfsubject={},
  pdfkeywords={},
  colorlinks=true,
  linkcolor=black,
  anchorcolor=black,
  citecolor=blue,
  urlcolor=blue,
}

<< "main-setup", echo = FALSE, warning=FALSE>>=
## setting knitr options (see https://yihui.org/knitr/ for all the options)
library(knitr)
opts_knit$set(concordance = FALSE)
opts_chunk$set(fig.height = 4,
               echo = FALSE,
               warning = FALSE,
               message = FALSE
               )

# Load necessary libraries
library(ggplot2)
library(cowplot)
library(gridExtra)
library(grid)
library(kableExtra)
library(latex2exp)
@

\begin{document}
\maketitle
\tableofcontents

\begin{abstract}
This report will focus on exploration of the paper ``A practical guide to Bayesian group sequential designs'' \citet{Gsponer}, and look at applications to paper ``Beyond p-values: A phase II dual-criterion design with statistical significance and clinical relevance'' \citet{Roychoudhury2018}.
\end{abstract}

\section{Background Knowledge}
\subsection{Dual Criteria Design}
%\begin{itemize}
%\item Statistical significance: The one-sided p-value must be less than $\alpha$. \item Clinical relevance: 
%\end{itemize}
By considering both the statistical significance and the clinical relevance criterion, we have both significant and sufficiently large effect estimate.\\
Required inputs: type I error control (null hypothesis (with null value ND) and type I error $\alpha$) and a decision value (DV). The DV is the ``target difference''. It is the minimal effect estimate needed for trail success (if higher than this value with moderate confidence, then a ``GO'' decision is made).\\
The dual-criterion is more demanding, the resulting power of study is less than that of a standard design.\\
For example, decisions for dual-criterion design when hazard ratio is the primary endpoint:
\begin{center}
<<plot1, echo=FALSE>>=
# Updated data based on visual inspection of the plot
data <- data.frame(
  group = factor(c("(1) NO-GO", "(2) GO", "(3)Inconclusive", "(4)Inconclusive"), 
                 levels = c("(4)Inconclusive", "(3)Inconclusive", 
                            "(2) GO","(1) NO-GO")),
  hazard_ratio = c(0.85, 0.6, 0.8, 0.65),     # Hazard ratio estimates for each case
  lower_ci = c(0.55, 0.25, 0.65, 0.25),         # Lower bounds of CIs
  upper_ci = c(1.15, 0.95, 0.95, 1.05)          # Upper bounds of CIs
)

# Plot setup
ggplot(data, aes(x = hazard_ratio, y = group)) +
  geom_point(shape = 18, size = 3) +   # Diamond marker for hazard ratios
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci), height = 0.2) + # Error bars for CIs
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +  # Line for HR=1
  geom_vline(xintercept = 0.7, linetype = "dashed", color = "grey50") + # Line for DV (assumed to be 0.7)
  annotate("text", x = 0.7, y = 0, label = "DV", vjust = -1, color = "red",fontface = "bold", size = 5) + # Label for DV
  annotate("text", x = 1.02, y = 0, label = "NV", vjust = -1, color = "red",fontface = "bold", size = 5) +   # Label for HR=1
  labs(x = "Hazard ratio", y = "") +
  scale_x_continuous(breaks = NULL, limits = c(0.2, 1.5)) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.minor = element_blank()) +
  annotate("text", x = 0.4, y = 4.5, label = "Treatment better", hjust = 0, color = "grey20") +
  annotate("text", x = 1.4, y = 4.5, label = "Treatment worse", hjust = 1, color = "grey20")
@
\end{center}
\textbf{Sample size calculation in dual criterion design}\\
Given the significance level $\alpha$, the null value (NV) and the decision value (DV), we can calculate the minimum sample size (for normally distributed data).
\begin{equation*}
n_{\text{min}}=\frac{\sigma^2 \times z_{\alpha}^2}{(\text{NV}-\text{DV})^2}
\end{equation*}
where $\sigma$ is the outcome standard deviation, and takes the value 2 under equal randomization for the standard normal approximation to time-to-event data. 
%In the log scale, if we say the value is within $\pm 2$ about the truth, then this translates into a range of approximately $\pm 100 \time 2 \%$ about the truth on the original scale. 
$z_{\alpha}$ is the $100(1-\alpha)\%$ quantile of the standard normal distribution. $n_{\text{min}}$ gives the minimum sample size that implies statistical significance if the effect estimate equals the DV. This value is calculated under the situation that both criterion are just satisfied. As illustrated in the below graph, when the effect estimate $\theta=$ DV, and the lower bound of the confidence interval just touches the NV so that statistical significance is reached, we have the minimum sample size. Notice that when sample size equals the minimum sample size, the half-width of the confidence interval $z_{\alpha}\sqrt{\frac{\sigma^2}{n_{\text{min}}}}$ equals NV-DV, so there will be no ``Inconclusive'' decisions. When the sample size is larger, the confidence interval becomes narrower, then an ``Inconclusive'' decision will occur.
\subsection{Group Sequential Designs}
``Group sequential'' means that the data analysis is conducted in interim analyses after every successive group of $2n$ patients. Fixing a maximum number of N groups, a trial is stopped at interim if the (two-sided) p-value is smaller than a pre-specified nominal significance level $\alpha$, or if N groups of patients have been recruited. The nominal significance level $\hat{\alpha}$ depends on the Type I error rate $\alpha$ and the number of groups N. Standard adjustments for multiple testing are too conservative, since tests are based on accumulating data with a specific dependence structure. \\
The commonly used stopping criteria for success and futility in Bayesian group sequential designs are related to treatment effect size and superior efficacy against placebo. They are often based on the posterior distribution of treatment effect at interim analyses. The Bayesian group sequential methods are designed to provide stopping criteria for related to subsequent decision making. \\
The operational characteristics of interest are probability to stop for success, to stop for futility, or to continue for each interim analysis and the final analysis, as well as corresponding cumulative probabilities. Also, the expected sample size is another characteristic to quantify the gain of group sequential design.

\subsection{Futility}
Futility refers to stopping a clinical trial early if interim results suggest that the treatment is unlikely to demonstrate a meaningful benefit by the end of the study. This helps to conserve resources, reduce unnecessary exposure to ineffective treatments, and allow quicker redirection of efforts to more promising therapies.
%There are two main types of futility stopping rules:
%	1.	Conditional Power-Based Futility – Stops the trial if the probability of achieving %statistical significance at the final analysis, given the interim data, is too low.
%	2.	Bounded Inferiority-Based Futility – Stops if the treatment effect is trending toward %being clinically insignificant or worse than the control.
\subsection{Interim analysis}
[to be added when we explore interim analysis]

\subsection{Mathematical Derivation of Bayesian Posterior Parameters} \label{Bayes}
We assume a normal prior distribution for the treatment effect \( \delta \): 
$\delta \sim N(\delta_0, \sigma_0^2)$, and a normal likelihood from observed data \( D_i \):
$D_i \sim N(\delta, V_i)$, where \( \delta_0 \) and \( \sigma_0^2 \) are the prior mean and variance, and \( V_i \) is the variance of the observed treatment effect.\\
Using Bayes' theorem:
$P(\delta \given D_i) \propto P(D_i \given \delta) P(\delta)$.\\
Expanding both normal densities in the exponent:
\begin{align*}
P(\delta \given D_i) &\propto \exp \left( -\frac{1}{2} \left[ \frac{(\delta - \delta_0)^2}{\sigma_0^2} + \frac{(D_i - \delta)^2}{V_i} \right] \right) \\
&\propto \exp \left( -\frac{1}{2} \left[ \left( \frac{1}{\sigma_0^2} + \frac{1}{V_i} \right) \delta^2 - 2 \left( \frac{\delta_0}{\sigma_0^2} + \frac{D_i}{V_i} \right) \delta \right] \right)
\end{align*}
Since a normal distribution has the general form:
$$\exp \left( -\frac{1}{2} \frac{(\delta - \mu_{\text{posterior}})^2}{\sigma_{\text{posterior}}^2} \right)=
\exp \left( -\frac{1}{2} \left[ \frac{1}{\sigma_{\text{posterior}}^2} \delta^2 - 2 \frac{\mu_{\text{posterior}}}{\sigma_{\text{posterior}}^2} \delta + C \right] \right),$$
comparing terms, the posterior mean and variance are:
$$\mu_{\text{posterior}} = \frac{\frac{\delta_0}{\sigma_0^2} + \frac{D_i}{V_i}}{\frac{1}{\sigma_0^2} + \frac{1}{V_i}}, \quad \quad \sigma_{\text{posterior}}^2 = \frac{1}{\frac{1}{\sigma_0^2} + \frac{1}{V_i}}.$$
This formula shows that the inverse variance (precision) of the posterior is the sum of the prior and data precisions. The posterior variance always smaller than either the prior or the likelihood variance.

\section{Methods}
\subsection{Effect and probability thresholds}
Traditionally, for treatment effect $\delta$, null hypothesis is $\delta \leq 0$, and alternative is $\delta > 0$. \\
Given type I error rate $\alpha$, power $1-\beta$, treatment effect size $\delta_1$ (minimal clinically relevant effect), in a standard two-group normal case with known standard deviation, the null will be rejected if observed average treatment effect $\hat{\delta}$ exceeds threshold $D_T=z_{1-\alpha} \cdot \frac{\delta_1}{z_{1-\alpha} + z_{1-\beta}}$. 
\begin{quote}
Deduce this $D_T$:
The null hypothesis is rejected when \( \hat{\delta} \) exceeds a threshold \( D_T \), so:
$P(\hat{\delta} > D_T \given \delta = 0) = \alpha$. Standardizing:
$P\left(Z > \frac{D_T}{\sigma/\sqrt{n}} \given \delta = 0 \right) = \alpha$,
which implies:
$\frac{D_T}{\sigma/\sqrt{n}} = z_{1-\alpha}$.
Thus, $D_T = z_{1-\alpha} \cdot \frac{\sigma}{\sqrt{n}}$ (1).\\
For the power condition (type II error control),
$P(\hat{\delta} > D_T \given \delta = \delta_1) = 1 - \beta$,
which gives:
$\frac{D_T - \delta_1}{\sigma/\sqrt{n}} = -z_{1-\beta}$. Rearranging, $D_T = \delta_1 - z_{1-\beta} \cdot \frac{\sigma}{\sqrt{n}}$ (2).\\
Equating (1) and (2), we get 
$\frac{\sigma}{\sqrt{n}} = \frac{\delta_1}{z_{1-\alpha} + z_{1-\beta}}$.\\
Substituting into the expression for \( D_T \):
$D_T = z_{1-\alpha} \cdot \frac{\delta_1}{z_{1-\alpha} + z_{1-\beta}}$. $\square$
\end{quote}
This threshold is typically smaller than $\delta_1$, so the null can be rejected for effect sizes that are not seen as clinically relevant.\\
To avoid situations where a statistically significant result is obtained with a small treatment effect, the double criterion is required to measure the success for clinical trials. \\
From Bayesian perspective, we base on posterior probabilities, and the success criteria are
\begin{enumerate}
\item $\mathbb{P}(\delta > s_1 \given \text{Data}) > p_1$,
\item $\mathbb{P}(\delta > s_2 \given \text{Data}) > p_2$,
\end{enumerate}
where $s_1, s_2$ are some effect thresholds, and $p_1, p_2$ are some probability thresholds.\\ 
Suppose posterior is normally distributed $\delta \sim \Nor(\mathbb{E}(\delta \given \text{Data}), \text{Sd}(\delta \given \text{Data}))$, then with minimal precision satisfying the double criteria, we have
$$\mathbb{E}(\delta \given \text{Data}) = \frac{s_1 z_{1-p_2} - s_2 z_{1-p_1}}{z_{1-p_2}-z_{1-p_1}}, \quad \text{Sd}(\delta \given \text{Data})=\frac{s_2-s_1}{z_{1-p_2}-z_{1-p_1}}.$$
\begin{quote}
\textbf{Derivation:}
The minimal precision means the maximal variance / widest confidence interval when the two criteria are just met. For larger precision, it may happen that one of the two criteria not satisfied. As shown in the below figure, the confidence interval in case (1) is too wide that the first criterion is not met. When at minimal precision (case (2)), two criteria are just met.
\begin{center}
<<minimal precision>>=
data <- data.frame(
  group = factor(c("(1) smaller precision than minimal", 
                   "(2) minimal precision", 
                   "(3) bigger precision than minimal"), 
                 levels = c("(3) bigger precision than minimal", 
                            "(2) minimal precision",
                            "(1) smaller precision than minimal")),
  s = c(1, 1, 1),     # Hazard ratio estimates for each case
  lower_ci = c(0.6, 0.7, 0.85),         # Lower bounds of CIs
  upper_ci = c(1.4, 1.3, 1.15)          # Upper bounds of CIs
)

# Plot setup
ggplot(data, aes(x = s, y = group)) +
  geom_point(shape = 18, size = 3) +   # Diamond marker for hazard ratios
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci), height = 0.2) + # Error bars for CIs
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +  # Line for HR=1
  geom_vline(xintercept = 0.7, linetype = "dashed", color = "grey50") + # Line for DV (assumed to be 0.7)
  annotate("text", x = 0.7, y = 0, label = "s1", vjust = -1, color = "red",fontface = "bold", size = 5) + 
  annotate("text", x = 1.02, y = 0, label = "s2", vjust = -1, color = "red",fontface = "bold", size = 5) +   
  labs(x = "effect thresholds", y = "") +
  scale_x_continuous(breaks = NULL, limits = c(0.2, 1.5)) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.minor = element_blank()) +
  annotate("text", x = 0.4, y = 3.5, label = "Treatment worse", hjust = 0, color = "grey20") +
  annotate("text", x = 1.4, y = 3.5, label = "Treatment better", hjust = 1, color = "grey20")
@
\end{center}

With the mean $\mathbb{E}(\delta \given \text{Data}) $ and standard deviation $\text{Sd}(\delta \given \text{Data})$, when the two criteria are just met:
\begin{align*}
    \mathbb{P}(\delta > s_1) &= 1 - \Phi\left(\frac{s_1 - \mathbb{E}(\delta \given \text{Data})}{\text{Sd}(\delta \given \text{Data})}\right) = p_1, \\
    \mathbb{P}(\delta > s_2) &= 1 - \Phi\left(\frac{s_2 - \mathbb{E}(\delta \given \text{Data})}{\text{Sd}(\delta \given \text{Data})}\right) = p_2.
\end{align*}
Because cdf is increasing:
\begin{align*}
    \frac{s_1 - \mathbb{E}(\delta \given \text{Data})}{\text{Sd}(\delta \given \text{Data})} &= z_{1 - p_1}, \\
    \frac{s_2 - \mathbb{E}(\delta \given \text{Data})}{\text{Sd}(\delta \given \text{Data})} &= z_{1 - p_2},
\end{align*}
Solving for expectation and standard deviatioin, 
$$
\mathbb{E}(\delta \given \text{Data}) = \frac{s_1 z_{1 - p_2} - s_2 z_{1 - p_1}}{z_{1 - p_2} - z_{1 - p_1}},$$
$$\text{Sd}(\delta \given \text{Data}) = \frac{s_2 - s_1}{z_{1 - p_2} - z_{1 - p_1}}. \quad \square$$
\end{quote}
Commonly, we use $s_1=0$, $s_2=\delta^*$, $p_1=1-\alpha$, and $p_2=0.5$, for the ease of interpretation and communication with clinicians. The $\delta^*$ is the required minimal observed difference for a trial to succeed. Then the success criteria are
\begin{enumerate}
\item $\mathbb{P}(\delta > 0 \given \text{Data}) > 1-\alpha$,
\item $\mathbb{P}(\delta > \delta^* \given \text{Data}) > 0.5$.
\end{enumerate}
This is similar to the double criterion in the frequentist approach when prior information is vague. Thus, the prior information for $\delta$ or for both treatment arms is crucial to Bayesian design.\\
We also define the criteria for futility:
\begin{enumerate}
\item $\mathbb{P}(\delta < f_1 \given \text{Data}) > q_1$,
\item $\mathbb{P}(\delta < f_2 \given \text{Data}) > q_2$.
\end{enumerate}
This is useful in the presence of interim analyses, because it could help to decide early stopping for futility. It is also useful in final analysis to classify a clear success, a clear failure, or an indeterminate outcome. The indeterminate case is undesirable, and happens when neither success criteria nor futility criteria are fulfilled.

\subsection{Interim analyses and operating characteristics}
Typically, MCMC methods are used for evaluating the posterior distribution. For evaluation of operating characteristics, people use approximate normalised likelihoods and normal priors.\\
For a two-arm clinical trial with one or more interim analyses, let $Y_{kij} \sim \Nor(\theta_k, \sigma^2)$ be the observation for treatment $k=1,2$, in stage $i=1,\ldots,l$, for subject $j=1,\ldots,n_{ki}$, with known standard deviation $\sigma$. The treatment effect is $\delta=\theta_2-\theta_1$. If prior information is $\delta \sim \Nor(\delta_0, \sigma_0^2)$, then the posterior at stage $i$ is 
\begin{equation*}
\delta \given \text{stage } i \sim \Nor\left( \omega_i \delta_0 + (1 - \omega_i) D_i, \frac{1}{\frac{1}{\sigma_0^2} + \frac{1}{V_i}} \right)
\end{equation*}
where $\omega_i = 1/(\sigma_0^2/V_i+1)$, $D_i$ is the aggregated treatment effect at stage $i$, and $V_i$ is the corresponding variance. Derivation of Bayesain posterior mean and variance could be found in Section \ref{Bayes}.\\
The term \( \omega_i \) represents the weight given to the prior mean in updating the posterior mean:
\begin{itemize}
\item If the prior variance \( \sigma_0^2 \) is large (weak prior), then \( \omega_i \approx 0 \), meaning the posterior is dominated by the observed data.
\item If the observed variance \( V_i \) is large (noisy data), then \( \omega_i \approx 1 \), meaning the posterior remains close to the prior.
\end{itemize}
If priors for two arms are available, say $\theta_1 \sim \Nor(\theta_{10},\sigma_1^2)$ and $\theta_2 \sim \Nor(\theta_{20},\sigma_2^2)$, then the posterior at stage $i$ is
\begin{equation*}
\delta \given \text{stage } i \sim \Nor\left( \{ \omega_{2i} \theta_{20} + (1 - \omega_{2i}) Y_{2i}\} - \{ \omega_{1i} \theta_{10} + (1 - \omega_{1i}) Y_{1i}\}, \frac{1}{\frac{1}{\sigma_1^2} + \frac{1}{V_{1i}}} + \frac{1}{\frac{1}{\sigma_2^2} + \frac{1}{V_{2i}}}\right),
\end{equation*}
where $Y_{ki}$ is the aggregated mean in arm $k$ in stage $i$, $V_{ki}$ is the corresponding variance, and $\omega_{ki}=1/(\sigma_k^2/V_{ki}+1)$.

\section{gsbDesign Package}

\subsection{Key functions in the package}
gsbDesign, gsbSimulation, gsb
<<gsbSimulation, echo=TRUE>>=
library(gsbDesign)
design <- gsbDesign(nr.stages = 2,
 patients = c(10,20),
 sigma=88,
 criteria.success = c(0,0.95,50,0.5),
 criteria.futility = c(40,0.9),
 prior.control = c(49,20),
 prior.treatment = c(49,0))
design
# simulation scenarios
simulation1 <- gsbSimulation(
  truth=list(49,c(0,40,50,60,70)),
  grid.type="sliced",
  method = "both",
  type.update="per arm",
  nr.sim=10000,
  warnings.sensitivity = 500,
  seed=1)
@

\noindent If \texttt{type.update = "treatment effect"}, the Bayesian update from prior to posterior is calculated on treatment effect delta. If \texttt{type.update = "per arm"}, the update is calculated separately in the placebo and the treatment arm. In this case it is possible to enter prior information in only one arm. For \texttt{type.update = "per arm"} only a simulation method is implemented.\\
If \texttt{type.update = "per arm"} and \texttt{grid.type = "table"} the argument has to be specified as list containing a vector of true control values and true treatment values.\\
If \texttt{type.update = "per arm"} and \texttt{grid.type = "sliced"} the argument has to be specified as list containing a vector of true control values and a vector of true deltas (= treatment - control).

\section{Reproducing results in Gsponer paper}
In this section we look at the four examples in the paper, to summarize and reproduce the results. In the next section, we are going to apply these methods to the cases in the paper by \cite{Roychoudhury2018}.

\subsection{Example 1: PoC trial on Crohn's disease} \label{Example1}
<< Example 1, echo=TRUE>>=
# design
library(gsbDesign)
design <- gsbDesign(nr.stages = 2,
 patients = c(10,20),
 sigma=88,
 criteria.success = c(0,0.95,50,0.5),
 criteria.futility = c(40,0.9),
 prior.control = c(49,20),
 prior.treatment = c(49,0))
design
# simulation scenarios
simulation <- gsbSimulation(
  truth=list(49,c(0,40,50,60,70)),
  grid.type="sliced",
  method = "both",
  type.update="per arm",
  nr.sim=10000,
  warnings.sensitivity = 500,
  seed=1)

#simulation1

# operating characteristics
result <- gsb(design=design,simulation=simulation)
#tables and graphics
tab(result, what="cumulative all")
tab(result, what="sample size")
plot(result, what="sample size", sliced=TRUE)
plot(result, what="cumulative all", sliced=TRUE)
@

\subsection{Example 2: Phase II single arm trials}
Simon's paper describes how we obtain the values in Table II. The response probability $p_0$=0.4 \cite{Simon}. In \citet{Simon}, they searched for optimal design (first stage sample size and total sample size) for a two-stage design.\\
Two relevant response rates: $\pi_C^*, \pi_E^*$. $\pi_C^*$ is the assumed historical control rate, and if true underlying response rate is larger than $\pi_E^*$, then it is likely that the drug is sufficiently promising. These correspond to $p_0, p_1$ in Simon's paper. $p_0$ is similar to the statistical significance criteria threshold, and $p_1$ is similar to the clinical relevance critaria threshold. In \citet{Gsponer}, the operating characteristics are obtained by package \citet{gsbDesign}, by reformulating single-arm design as two-arm design, and using normal approximation for the endpoint. For $r$ responders in $n$ patients, define $y$ as log-odds of observed response rate. Its distribution is approximately 
$$y \sim \Nor(\logit(\pi_E), \sigma_E^2/n),$$
where $\pi_E$ is the true response rate, and $\sigma_E^2=1/\pi_E + 1/(1-\pi_E)$.

\begin{quote}
Derivation of approximated variance: $\sigma_E^2/n$\\
Logit function $g(\pi) = \log \left( \frac{\pi}{1 - \pi} \right)$ has derivative $g'(\pi) = \frac{1}{\pi(1 - \pi)}$.\\ 
Applying the Delta Method:
$\text{Var}(\logit(\pi)) \approx \left( \frac{1}{\pi(1 - \pi)} \right)^2 \text{Var}(\pi).$\\
Since $\text{Var}(\pi) \approx \frac{\pi (1 - \pi)}{n}$, we get 
$\text{Var}(\logit(\pi)) \approx \frac{1}{\pi(1 - \pi)} \cdot \frac{\pi(1 - \pi)}{n} = \frac{1}{n \pi(1 - \pi)}. \quad \square$ 
\end{quote}

A virtual control arm is needed for this single-arm design to mimic a two-arm design. This is done by setting sample size in control group to 0, and by providing an extremely informative prior. The prior is chosen to be centered at $\pi_C^*$ with an extremely small variance.
<<Example 2, echo=TRUE>>=
logit <- function(p) { log(p/(1-p)) }
expit <- function(x) { exp(x)/(1+exp(x)) }
p0 = 0.4
p1 = 0.6
suc = logit(23.5/46) - logit(p0)
fut = logit(7.5/16) - logit(p0)
design <- gsbDesign(nr.stages=2,
 patients=cbind( c(0,0), c(16,30) ),
 sigma=2,
 criteria.success=rbind( c(NA,NA), c(suc,0.5) ),
 criteria.futility=rbind( c(fut,0.5), c(NA,NA) ),
 prior.control=c( logit(p0),1000),
 prior.treatment=c( logit(p0),0.001))

simulation <- gsbSimulation(truth=list(logit(p0),seq(0,0.8,0.01)),
 grid.type = "sliced",
 type.update = "per arm",
 nr.sim = 100000,
 warnings.sensitivity = 2000,
 seed = 1)
result <- gsb(design,simulation)
result
#tab(result, what="cumulative all")
#tab(result, what="sample size")
plot(result, what="sample size", sliced=TRUE)
plot(result, what="cumulative all", sliced=TRUE)
par(mfrow=c(1,2))
y1 = tab(result,what='futility')[,4]
y2 = tab(result,what='success')[,5]
y3 = 1 - y1 - y2
piE = expit( -0.4+seq(0,0.8,0.01) ) # transform back to original scale
plot(piE,y1,type='l',ylim=c(0,1),ylab='probability') 
lines(piE,y2)
lines(piE,y3)
y3 = tab(result,what='sample size')[,5]
plot(piE,y3,type='l',ylab='expected sample size') 
@

\subsection{Example 3: Phase II trial with count endpoint}
<<Example 3 case 1, echo=TRUE>>=
### Case 1 non-informative prior
kappa=2 ; rate1=log(14.8) ; redscen=seq(0,0.9,0.05); OCout2=NULL
for(i in seq(along=redscen)){
red=redscen[i]
rate2=log(14.8*(1-red))
delta=rate1-rate2
sigma1=sqrt(kappa+(1/rate1)) ; sigma2=sqrt(kappa+(1/rate2))
design1 <- gsbDesign(nr.stages=2,
patients=25,
sigma=c(sigma2,sigma1),
criteria.success=c(0,0.9, log(2.5), 0.5),
criteria.futility=c(log(1.42),0.5,log(2.5),0.9),
 prior.difference="non-informative")
simulation1 <- gsbSimulation(truth=c(-1,delta,2), type.update="treatment effect",
 method= "numerical integration")
x1 <- gsb(design=design1, simulation=simulation1)
sub=x1$OC[,"delta"]!=-1
xout=x1$OC[sub,]
xout$delta=red*100
OCout2=rbind(OCout2,xout)
}
x1$OC=OCout2
p1=plot(x1,what="cumulative all")
p1$ylab="Cumulative probability of success or futility"; p1$xlab="% relative reduction\n"
p1$main=NULL
p1$panel.args.common$col <- 1 ; p1$panel.args.common$lty <- 1:3
p1$legend$bottom$args$key$lines$col <- 1 ; p1$legend$bottom$args$key$lines$lty <- 1:3
plot(p1) 
@

<<Example 3 case 2, echo=TRUE>>=
### Case 2 -informative prior
kappa=2 ; rate1=log(14.8); redscen=seq(0,0.9,0.05) ; OCout=NULL
for(i in seq(along=redscen)){
red=redscen[i]
rate2=log(14.8*(1-red))
delta=rate1-rate2
delta.prior=rate1-log(14.8*(1-0.75))
sigma1=sqrt(kappa+(1/rate1)) ; sigma2=sqrt(kappa+(1/rate2))
design1 <- gsbDesign(nr.stages=2,
patients=25,sigma=c(sigma2,sigma1),
criteria.success=c(0,0.9, log(2.5), 0.5),
criteria.futility=c(log(1.42),0.5,log(2.5),0.9),
 prior.difference=c(delta.prior,10,10))
simulation1 <- gsbSimulation(truth=c(-1,delta,2),
 type.update="treatment effect",
 method= "numerical integration")
x1 <- gsb(design=design1, simulation=simulation1)
sub=x1$OC[,"delta"]!=-1
xout=x1$OC[sub,]
xout$delta=red*100
OCout=rbind(OCout,xout)
}
x1$OC=OCout
p1=plot(x1,what="cumulative all")
p1$xlab="% relative reduction\n" ; p1$ylab="Cumulative probability of success or futility"
p1$main=NULL
p1$panel.args.common$col <- 1 ; p1$panel.args.common$lty <- 1:3
p1$legend$bottom$args$key$lines$col <- 1 ; p1$legend$bottom$args$key$lines$lty <- 1:3
plot(p1) 
@

\subsection{Example 4: Phase III trial with Bayesian futility criteria for time-to-event endpoint}
<<Example 4, echo=TRUE>>=
succ <- rbind(c(0,0.99995),
 c(0,0.9995),
 c(log(1.075), 0.975))
fut <- rbind(c(0,0.99995),
 c(0,0.9995),
 NA)
## additional rules for stopping for futility (alternative design)
fut1 <- rbind(c(log(1.075),0.95),
 c(log(1.075),0.9),
 c(log(1.075),0.8))
## exchange success and futility criteria as smaller values of the endpoint are better here
design1 <- gsbDesign(nr.stages = 3,
 patients = 300,
 criteria.success = fut,
 criteria.futility = succ,
 sigma = 1,
 prior.difference = "non-informative")
design2 <- gsbDesign(nr.stages = 3,
 patients = 300,
 criteria.success = fut1,
 criteria.futility = succ,
 sigma = 1,
 prior.difference = "non-informative")
truth <- seq(-0.15,0.15,by=0.005)
sim <- gsbSimulation(truth = truth,
 grid.type = "manually",
 type.update = "treatment effect",
 method = "numerical integration")

## Calculate the operating characteristics
x1 <- gsb(design = design1, simulation = sim)
x2 <- gsb(design = design2, simulation = sim)
## plot of outcome probabilities (on original HR scale (not log-scale))
dep <- subset(x1$OC,
 x1$OC$type %in% c("cumulative futility", "cumulative success"))
dep$stage <- paste("actual", dep$stage, sep=", ")
dep2 <- subset(x2$OC,
 x2$OC$type %in% c("cumulative futility", "cumulative success"))
dep2$stage <- paste("alternative futility", dep2$stage, sep=", ")
dep0 <- rbind(dep, dep2)
p1 <- xyplot(value~exp(delta)|stage, groups=type, data=dep0, col =1, lty=c(2,2,1),
 layout = c(3,2), as.table = TRUE, scales = list(alternating = 1),
 ylab = "Cumulative Probability of Stopping",
 ylim = c(-0.05,1.05), xlab = "True Hazard Ratio\n",
 panel = function(...){
 panel.grid(h=-1,v=-1, col="grey", lwd=1, lty=2)
 panel.xyplot(... , type="l", lwd=3)},
 key = list(columns = 2, space = "bottom", lines = list(lty=1:2, lwd = 2.5),
 text = list(c("Futility", "Success" ), col = 1), border = FALSE))
plot(p1)
## plot for expected sample size (on original HR scale (not log-scale))
xx1 <- subset(x1$OC, x1$OC$type %in% c("sample size"))
xx2 <- subset(x2$OC, x2$OC$type %in% c("sample size"))
xx <- rbind(cbind(xx1, design = "actual design"),
 cbind(xx2, design = "alternative design"))
xx <- subset(xx, xx$stage %in% c("stage 3"))
xx$value <- xx$value*1/0.09

p2 <- xyplot(value~exp(delta)|design, data=xx,
 scales = list(alternating = 1),
 panel = function(...){
 panel.grid(h=-1,v=-1, col="grey", lwd=1, lty=2)
 panel.xyplot(... , type="l", lwd=3, col=1)
 }, ylim = c(1300*1/0.09, 1850*1/0.09),
 ylab = "Expected Sample Size", xlab = "True Hazard Ratio")
plot(p2)
@

\section{Likelihood Ratios, Unit Information LR, and DOR}
Likelihood ratios (LR) quantify the change in disease odds after having
observed a positive or negative test result. LRs depend on power and Type I error:
$$\text{LR}^+ = \frac{\text{Power}}{\text{Type I Error}}, \quad \text{LR}^- = \frac{1-\text{Power}}{1-\text{Type I Error}}.$$
In Example 1, we have the following table of operating characteristics of the two-stage design for the PoC trial in Crohn's disease. The probability of success when true $\delta=0$ (no difference between placebo and treatment) is the Type I error. The probabilities of success when true $\delta=40, 50, 60, 70$ are the powers in each case. It would be interesting to look at how the likelihood ratios and relevant statistics change with the effect size $\delta$.
<<Example 1 table, echo=FALSE, results='asis'>>=
library(gsbDesign)
design <- gsbDesign(nr.stages = 2,
 patients = c(10,20),
 sigma=88,
 criteria.success = c(0,0.95,50,0.5),
 criteria.futility = c(40,0.9),
 prior.control = c(49,20),
 prior.treatment = c(49,0))
# simulation scenarios
simulation <- gsbSimulation(
  truth=list(49,c(0,40,50,60,70)),
  grid.type="sliced",
  method = "both",
  type.update="per arm",
  nr.sim=100000,
  warnings.sensitivity = 500,
  seed=1)

# operating characteristics
result <- gsb(design=design,simulation=simulation)
#tables and graphics
table <- as.matrix(tab(result, what="cumulative all")[,c(3,7:9)])
table <- cbind(table, tab(result, what="sample size")[,5])
colnames(table) <- c("$\\delta$", "Success",
                         "Futile","Indeterminate","Expected N")

kable(table, format = "latex", booktabs = TRUE, escape = FALSE,
      caption = "Operating Charactristics") %>%
  kable_styling(latex_options = c("hold_position", "striped")) %>%
  cat()
@
\pagebreak
\subsection{Positive and Negative Likelihood Ratios}
\begin{center}
<<LRs, echo=TRUE>>=
library(latex2exp)
delta <- c(40,50,60,70) # true effect sizes
Exp_N <- table[,5]
TypeIError <- table[1,2]
power <- table[2:5,2]
LR_positive <- power/TypeIError
LR_negative <- (1-power)/(1-TypeIError)

plot(delta, LR_positive, type="o", col="blue", lwd=2, xaxt="n", ylim = c(30,83),
     xlab=TeX(r"($\delta$)"), ylab="Positive Likelihood Ratio")
axis(1, at=c(40, 50, 60, 70))
# Add horizontal text labels on the y-axis
text(delta-1.3, LR_positive+3, labels=round(LR_positive,2), pos=4, cex=0.8, col="red")

plot(delta, LR_negative, type="o", col="blue", lwd=2, xaxt="n", ylim = c(0.08,0.68),
     xlab=TeX(r"($\delta$)"), ylab="Negative Likelihood Ratio")
axis(1, at=c(40, 50, 60, 70))
# Add horizontal text labels on the y-axis
text(delta-1, LR_negative+0.05, labels=round(LR_negative,2), pos=4, cex=0.8, col="red")

@
\end{center}
\subsection{Unit Information Likelihood Ratios}
When we receive multiple independent pieces of evidence  $E_1, E_2, …, E_n$, the total likelihood ratio is multiplicative:
$$\text{LR}_{\text{total}} = \text{LR}_1 \times \text{LR}_2 \times … \times \text{LR}_n$$
Taking the n$^{th}$ root, the Unit Information Likelihood Ratio (LR$_{\text{UI}}$) is:
$$\text{LR}_{\text{UI}} = (\text{LR}_{\text{total}})^{1/n}.$$
This tells us how much each piece of independent observation contributes on average to the total likelihood ratio.\\
We know $$\text{Posterior odds} =
\left\{
\begin{array}{l}
\text{LR}^{+} = \left\{ \text{LR}_{\text{UI}}^{+} \right\}^{\mathbb{E}(N)} \\
\text{LR}^{-} = \left\{ \text{LR}_{\text{UI}}^{-} \right\}^{\mathbb{E}(N)}
\end{array}
\right\}
\cdot \text{Prior odds}.$$
This enables Bayesian updating with likelihood ratios, as likelihood ratios have a multiplicative nature. Then, we can also obtain unit information likelihood ratios ($\text{LR}_{\text{UI}}=\text{LR}^{1/\mathbb{E}(N)}$), which is the likelihood ratio per unit given the expected sample size. This tells us the information one observation contributes.\\
For each effect size $\delta$, we can compute two bounds for their UI LRs based on the expected sample size $\mathbb{E}(N)$ for that particular value of $\delta$ used to calculate the power and for $\delta=0$. 
\begin{center}
<<Unit information, echo=TRUE>>=
LR_pos_UI <- as.data.frame(t(matrix(
  c(max(LR_positive[1]^(1/Exp_N[1:2])),min(LR_positive[1]^(1/Exp_N[1:2])),
    max(LR_positive[2]^(1/Exp_N[c(1,3)])), min(LR_positive[2]^(1/Exp_N[c(1,3)])),
    max(LR_positive[3]^(1/Exp_N[c(1,4)])), min(LR_positive[3]^(1/Exp_N[c(1,4)])),
    max(LR_positive[4]^(1/Exp_N[c(1,5)])), min(LR_positive[4]^(1/Exp_N[c(1,5)]))),
  nrow = 2, ncol = 4,
  dimnames = list(c("upper bound", "lower bound"),
                  c("delta=40","delta=50","delta=60","delta=70"))
  )))

LR_neg_UI <- as.data.frame(t(matrix(
  c(max(LR_negative[1]^(1/Exp_N[1:2])),min(LR_negative[1]^(1/Exp_N[1:2])),
    max(LR_negative[2]^(1/Exp_N[c(1,3)])), min(LR_negative[2]^(1/Exp_N[c(1,3)])),
    max(LR_negative[3]^(1/Exp_N[c(1,4)])), min(LR_negative[3]^(1/Exp_N[c(1,4)])),
    max(LR_negative[4]^(1/Exp_N[c(1,5)])), min(LR_negative[4]^(1/Exp_N[c(1,5)]))),
  nrow = 2, ncol = 4,
  dimnames = list(c("upper bound", "lower bound"),
                  c("delta=40","delta=50","delta=60","delta=70"))
  )))

# Plot UI LR+
plot(delta, LR_pos_UI$`upper bound`, type = "n", xaxt="n",
     ylim = range(LR_pos_UI$`lower bound`,LR_pos_UI$`upper bound`),
     xlab = TeX(r"($\delta$)"), ylab = TeX(r"(Unit information LR$^+$)"), 
     main = TeX(r"(UI LR$^+$ bounds)"))
axis(1, at=c(40, 50, 60, 70))

# Shade the area between bounds
polygon(c(delta, rev(delta)), 
        c(LR_pos_UI$`upper bound`, rev(LR_pos_UI$`lower bound`)), 
        col = rgb(0.2, 0.5, 0.8, 0.3), border = NA)

# Add lines for upper and lower bounds
lines(delta, LR_pos_UI$`upper bound`, col = "blue3", lwd = 2)
lines(delta, LR_pos_UI$`lower bound`, col = "red3", lwd = 2)

# Plot UI LR-
plot(delta, LR_neg_UI$`upper bound`, type = "n", xaxt="n",
     ylim = range(LR_neg_UI$`lower bound`,LR_neg_UI$`upper bound`),
     xlab = TeX(r"($\delta$)"), ylab = TeX(r"(Unit information LR$^-$)"), 
     main = TeX(r"(UI LR$^-$ bounds)"))
axis(1, at=c(40, 50, 60, 70))

# Shade the area between bounds
polygon(c(delta, rev(delta)), 
        c(LR_neg_UI$`upper bound`, rev(LR_neg_UI$`lower bound`)), 
        col = rgb(0.2, 0.5, 0.8, 0.3), border = NA)

# Add lines for upper and lower bounds
lines(delta, LR_neg_UI$`upper bound`, col = "blue3", lwd = 2)
lines(delta, LR_neg_UI$`lower bound`, col = "red3", lwd = 2)

@
\end{center}

\subsection{Diagnostic Odds Ratios}
Diagnostic odds ratios (DOR) tells us how much information there is essentially. The unit information DOR implies the amount of information provided by one observation.
\begin{center}
<<Diagnostic OR, echo=TRUE>>=
DOR <- LR_positive/LR_negative
A <- as.matrix(LR_pos_UI)
B <- as.matrix(LR_neg_UI)
DOR_UI <- t(sapply(1:nrow(LR_pos_UI), function(i) as.vector(outer(A[i, ], B[i, ], "/"))))
DOR_UI <- as.data.frame(cbind(DOR_UI, upper_bound=apply(DOR_UI,1, max), 
                              lower_bound=apply(DOR_UI,1, min)))

# Plot DOR
plot(delta, DOR, type = "o", xaxt="n",lty=1,lwd=2,
     ylim = range(DOR),
     xlab = TeX(r"($\delta$)"), ylab = "DOR", col="blue",
     main = TeX(r"(DOR)"))
axis(1, at=c(40, 50, 60, 70))

# Plot DOR UI
plot(delta, DOR_UI$upper_bound, type = "n", xaxt="n",
     ylim = range(DOR_UI$lower_bound,DOR_UI$upper_bound),
     xlab = TeX(r"($\delta$)"), ylab = TeX(r"(Unit information DOR)"), 
     main = TeX(r"(UI DOR bounds)"))
axis(1, at=c(40, 50, 60, 70))

# Shade the area between bounds
polygon(c(delta, rev(delta)), 
        c(DOR_UI$upper_bound, rev(DOR_UI$lower_bound)), 
        col = rgb(0.2, 0.5, 0.8, 0.3), border = NA)

# Add lines for upper and lower bounds
lines(delta, DOR_UI$upper_bound, col = "blue3", lwd = 2)
lines(delta, DOR_UI$lower_bound, col = "red3", lwd = 2)

@
\end{center}

\subsection{Exploring the impact of the presence of decision value}
Here we remove the DV criterion in Example 1 and see how the results change. Below table contains the operating characteristics and expected sample size.
<<Remove DV, echo=FALSE, results='asis'>>=
design <- gsbDesign(nr.stages = 2,
 patients = c(10,20),
 sigma=88,
 criteria.success = c(0,0.95),
 criteria.futility = c(40,0.9),
 prior.control = c(49,20),
 prior.treatment = c(49,0))

simulation <- gsbSimulation(
  truth=list(49,c(0,40,50,60,70)),
  grid.type="sliced",
  method = "both",
  type.update="per arm",
  nr.sim=100000,
  warnings.sensitivity = 500,
  seed=1)

# operating characteristics
result_noDV <- gsb(design=design,simulation=simulation)
#tables and graphics
table_noDV <- as.matrix(tab(result_noDV, what="cumulative all")[,c(3,7:9)])
table_noDV <- cbind(table_noDV, tab(result_noDV, what="sample size")[,5])
colnames(table_noDV) <- c("$\\delta$", "Success",
                         "Futile","Indeterminate","Expected N")

kable(table_noDV, format = "latex", booktabs = TRUE, escape = FALSE,
      caption = "Probability overall") %>%
  kable_styling(latex_options = c("hold_position", "striped")) %>%
  cat()
@
\\
Again we look at the LR$^+$, LR$^-$, unit information LR$_{\text{UI}}^+$, LR$_{\text{UI}}^-$, DOR and DOR$_{\text{UI}}$. Generally, LR$^+$, LR$^-$, LR$_{\text{UI}}^+$, and LR$_{\text{UI}}^+$ are smaller that that under the presence of DV criterion, while the DOR$_{\text{UI}}$s are generally larger, which could be seen from the below plots.
\begin{center}
<<LR and DOR without DV, echo=FALSE>>=
delta <- c(40,50,60,70) # true effect sizes
Exp_N_noDV <- table_noDV[,5]
TypeIError_noDV <- table_noDV[1,2]
power_noDV <- table_noDV[2:5,2]

# LR+ and LR-
LR_positive_noDV <- power_noDV/TypeIError_noDV
LR_negative_noDV <- (1-power_noDV)/(1-TypeIError_noDV)

# plot LR+ with and without DV
plot(delta, LR_positive_noDV, type="o", col="blue", lty=2,
     lwd=2, xaxt="n", ylim =range(LR_positive_noDV,LR_positive+3),
     xlab=TeX(r"($\delta$)"), ylab="Positive Likelihood Ratio")
axis(1, at=c(40, 50, 60, 70))
lines(delta, LR_positive, type="o",col="blue",lwd=2)
text(delta-1.3, LR_positive_noDV+2, labels=round(LR_positive_noDV,2), 
     pos=4, cex=0.8, col="red")
text(delta-1.3, LR_positive+2, labels=round(LR_positive,2), pos=4, cex=0.8, col="red")
legend("topleft", legend = c("with DV","without DV"),
       col = c("blue", "blue"),
       lty = c(1,2), lwd = c(2,2),bty = "n")

# plot LR- with and without DV
plot(delta, LR_negative_noDV, type="o", col="blue", lty=2,
     lwd=2, xaxt="n", ylim =range(LR_negative_noDV,LR_negative+0.5),
     xlab=TeX(r"($\delta$)"), ylab="Negative Likelihood Ratio")
axis(1, at=c(40, 50, 60, 70))
lines(delta, LR_negative, type="o",col="blue",lwd=2)
text(delta-1, LR_negative_noDV+0.04, labels=round(LR_negative_noDV,2), 
     pos=4, cex=0.8, col="red")
text(delta-1, LR_negative+0.04, labels=round(LR_negative,2), pos=4, cex=0.8, col="red")
legend("topright", legend = c("with DV","without DV"),
       col = c("blue", "blue"),
       lty = c(1,2), lwd = c(2,2),bty = "n")

# UI LR:
LR_pos_UI_noDV <- as.data.frame(t(matrix(
  c(max(LR_positive_noDV[1]^(1/Exp_N_noDV[1:2])),min(LR_positive_noDV[1]^(1/Exp_N_noDV[1:2])),
    max(LR_positive_noDV[2]^(1/Exp_N_noDV[c(1,3)])), min(LR_positive_noDV[2]^(1/Exp_N_noDV[c(1,3)])),
    max(LR_positive_noDV[3]^(1/Exp_N_noDV[c(1,4)])), min(LR_positive_noDV[3]^(1/Exp_N_noDV[c(1,4)])),
    max(LR_positive_noDV[4]^(1/Exp_N_noDV[c(1,5)])), min(LR_positive_noDV[4]^(1/Exp_N_noDV[c(1,5)]))),
  nrow = 2, ncol = 4,
  dimnames = list(c("upper bound", "lower bound"),
                  c("delta=40","delta=50","delta=60","delta=70"))
  )))

LR_neg_UI_noDV <- as.data.frame(t(matrix(
  c(max(LR_negative_noDV[1]^(1/Exp_N_noDV[1:2])),min(LR_negative_noDV[1]^(1/Exp_N_noDV[1:2])),
    max(LR_negative_noDV[2]^(1/Exp_N_noDV[c(1,3)])), min(LR_negative_noDV[2]^(1/Exp_N_noDV[c(1,3)])),
    max(LR_negative_noDV[3]^(1/Exp_N_noDV[c(1,4)])), min(LR_negative_noDV[3]^(1/Exp_N_noDV[c(1,4)])),
    max(LR_negative_noDV[4]^(1/Exp_N_noDV[c(1,5)])), min(LR_negative_noDV[4]^(1/Exp_N_noDV[c(1,5)]))),
  nrow = 2, ncol = 4,
  dimnames = list(c("upper bound", "lower bound"),
                  c("delta=40","delta=50","delta=60","delta=70"))
  )))

# Plot UI LR+ with and without DV
plot(delta, LR_pos_UI_noDV$`upper bound`, type = "n", xaxt="n",
     ylim = range(LR_pos_UI_noDV$`lower bound`,LR_pos_UI$`upper bound`),
     xlab = TeX(r"($\delta$)"), ylab = TeX(r"(Unit information LR$^+$)"), 
     main = TeX(r"(UI LR$^+$ bounds)"))
axis(1, at=c(40, 50, 60, 70))

# Shade the area between bounds
polygon(c(delta, rev(delta)), 
        c(LR_pos_UI_noDV$`upper bound`, rev(LR_pos_UI_noDV$`lower bound`)), 
        col = rgb(1, 0.6, 0.2, 0.3), border = NA)

# Add lines for upper and lower bounds
lines(delta, LR_pos_UI_noDV$`upper bound`, col = "blue3", lwd = 2, lty = 2)
lines(delta, LR_pos_UI_noDV$`lower bound`, col = "red3", lwd = 2, lty = 2)

# add bounds when there is DV
polygon(c(delta, rev(delta)), 
        c(LR_pos_UI$`upper bound`, rev(LR_pos_UI$`lower bound`)), 
        col = rgb(0.2, 0.5, 0.8, 0.3), border = NA)

# Add lines for upper and lower bounds
lines(delta, LR_pos_UI$`upper bound`, col = "blue3", lwd = 2)
lines(delta, LR_pos_UI$`lower bound`, col = "red3", lwd = 2)

legend("topleft", legend = c("with DV", NA,
                              "without DV", NA),
       col = c("blue3", "red3", "blue3", "red3"),
       lty = c(1, 1, 2, 2), lwd = c(2, 2, 2, 2),bty = "n")

# Plot UI LR- with and without DV
plot(delta, LR_neg_UI_noDV$`upper bound`, type = "n", xaxt="n",
     ylim = range(LR_neg_UI_noDV$`lower bound`,LR_neg_UI$`upper bound`),
     xlab = TeX(r"($\delta$)"), ylab = TeX(r"(Unit information LR$^-$)"), 
     main = TeX(r"(UI LR$^-$ bounds)"))
axis(1, at=c(40, 50, 60, 70))

# Shade the area between bounds
polygon(c(delta, rev(delta)), 
        c(LR_neg_UI_noDV$`upper bound`, rev(LR_neg_UI_noDV$`lower bound`)), 
        col = rgb(1, 0.6, 0.2, 0.3), border = NA)

# Add lines for upper and lower bounds
lines(delta, LR_neg_UI_noDV$`upper bound`, col = "blue3", lwd = 2,lty = 2)
lines(delta, LR_neg_UI_noDV$`lower bound`, col = "red3", lwd = 2, lty = 2)

# add bounds when there is DV
polygon(c(delta, rev(delta)), 
        c(LR_neg_UI$`upper bound`, rev(LR_neg_UI$`lower bound`)), 
        col = rgb(0.2, 0.5, 0.8, 0.3), border = NA)

# Add lines for upper and lower bounds
lines(delta, LR_neg_UI$`upper bound`, col = "blue3", lwd = 2)
lines(delta, LR_neg_UI$`lower bound`, col = "red3", lwd = 2)

legend("bottomleft", legend = c("with DV", NA,
                                "without DV",NA),
       col = c("blue3", "red3", "blue3", "red3"),
       lty = c(1, 1, 2, 2), lwd = c(2, 2, 2, 2),bty = "n")

# DOR 
DOR_noDV <- LR_positive_noDV/LR_negative_noDV
A <- as.matrix(LR_pos_UI_noDV)
B <- as.matrix(LR_neg_UI_noDV)
DOR_UI_noDV <- t(sapply(1:nrow(A), function(i) as.vector(outer(A[i, ], B[i, ], "/"))))
DOR_UI_noDV <- as.data.frame(cbind(DOR_UI_noDV, upper_bound=apply(DOR_UI_noDV,1, max), 
                              lower_bound=apply(DOR_UI_noDV,1, min)))

# Plot DOR with and without DV
plot(delta, DOR_noDV, type = "o", xaxt="n",lty=2,lwd=2,
     ylim = range(DOR_noDV, DOR),
     xlab = TeX(r"($\delta$)"), ylab = "DOR", col="blue",
     main = TeX(r"(DOR)"))
axis(1, at=c(40, 50, 60, 70))
lines(delta, DOR, type="o", col="blue",lty=1, lwd=2)
legend("topleft", legend = c("with DV","without DV"),
       col = c("blue", "blue"),
       lty = c(1,2), lwd = c(2,2),bty = "n")

# Plot DOR UI with and without DV
plot(delta, DOR_UI_noDV$upper_bound, type = "n", xaxt="n",
     ylim = range(DOR_UI$lower_bound,DOR_UI_noDV$upper_bound),
     xlab = TeX(r"($\delta$)"), ylab = TeX(r"(Unit information DOR)"), 
     main = TeX(r"(UI DOR bounds)"))
axis(1, at=c(40, 50, 60, 70))

# Shade the area between bounds
polygon(c(delta, rev(delta)), 
        c(DOR_UI_noDV$upper_bound, rev(DOR_UI_noDV$lower_bound)), 
        col = rgb(1, 0.6, 0.2, 0.3), border = NA)

# Add lines for upper and lower bounds
lines(delta, DOR_UI_noDV$upper_bound, col = "blue3", lwd = 2, lty = 2)
lines(delta, DOR_UI_noDV$lower_bound, col = "red3", lwd = 2, lty = 2)

# add bounds when there is DV
polygon(c(delta, rev(delta)), 
        c(DOR_UI$upper_bound, rev(DOR_UI$lower_bound)), 
        col = rgb(0.2, 0.5, 0.8, 0.3), border = NA)

# Add lines for upper and lower bounds
lines(delta, DOR_UI$upper_bound, col = "blue3", lwd = 2)
lines(delta, DOR_UI$lower_bound, col = "red3", lwd = 2)

legend("topleft", legend = c("with DV", NA, "without DV", NA),
       col = c("blue3", "red3", "blue3", "red3"),
       lty = c(1, 1, 2, 2), lwd = c(2, 2, 2, 2),bty = "n")
@
\end{center}

\section{Applications to Roychoudhury paper}
In this part, we look at the application of the methods in \citet{Gsponer} to the 
single-arm PoC design with binary data in \citet{Roychoudhury2018}. \\
For binary data, the normal approximation is known to be more appropriate on the logit scale than on the proportion scale. Let $r$ be number of responders, and $n$ be the number of treated patients. Defining y as the corresponding log-odds of the observed response rate, its distribution is approximately given by 
\begin{equation*}
y \sim \Nor(\logit(\pi),\sigma^2/n)
\end{equation*}
where $\pi$ is the true response rate and $\sigma^2=1/\pi+1/(1-\pi)$. \\
In \citet{Roychoudhury2018} paper, the primary endpoint is objective response rate(ORR). The prior is set to be ORR $\sim$ Beta(0.0811, 1). Because of the absence of a comparator (in single arm trials), NV is set to 7.5\%. A minimum improvement of 10\% is considered necessary for further development, so the DV is set to be 10\%+7.5\%=17.5\%. Notice in this example, NV $<$ DV.\\
The dual-criterion is:
\begin{enumerate}
\item Bayesian statistical significance: $\mathbb{P}(\text{ORR} \geq 7.5\% \given \text{data}) \geq 0.95$
\item Clinical relevance: Posterior median $\geq 17.5\%$
\end{enumerate}
Null hypothesis: there is no effect of the drug, i.e. ORR=7.5\%\\
If we observe a number of successes ($r$) and failures ($n - r$), then the posterior is Beta($a + r$, $b + n - r$).\\
One key difference between \citet{Gsponer} and \citet{Roychoudhury2018} is that \citet{Gsponer} is using a logit transformation for normal approximation, whereas \citet{Roychoudhury2018} is using a Beta prior. So here we need to translate \citet{Roychoudhury2018}'s problem into logit scale:\\
The prior is approximately logit(ORR) $\sim \Nor(-12,12^2)$. This can be visualized below in the histogram of logit(ORR) where ORR $\sim$ Beta(0.0811, 1), and the red curve is the density of $\Nor(-12, 12^2)$.
\begin{center}
<<normal approximation to Beta, echo=FALSE, results='hide'>>=
ORR <- rbeta(10000, 0.0811, 1)
logitORR <- log(ORR/(1-ORR))
mean(logitORR)
sd(logitORR)
hist(logitORR, probability = TRUE, xlim = c(-80,15),breaks = 30,
     main = "Normal approximation to logitORR")
x = seq(-80,10)
curve(dnorm(x, mean = -12, sd = 12), xlim = c(-80,15),
      col = "red", lwd = 2, add = TRUE)
@
\end{center}
The translated dual-criterion for success are:
\begin{enumerate}
\item Bayesian statistical significance: $\mathbb{P}(\text{logit(ORR)} \geq \logit(0.075) \given \text{data}) \geq 0.95$
\item Clinical relevance: $\mathbb{P}(\text{logit(ORR)} \geq \logit(0.175) \given \text{data}) \geq 0.5$
\end{enumerate}
The translated dual-criterion for failure are:
\begin{enumerate}
\item Bayesian statistical significance: $\mathbb{P}(\text{logit(ORR)} \leq \logit(0.075) \given \text{data}) \geq 0.05$
\item Clinical relevance: $\mathbb{P}(\text{logit(ORR)} \leq \logit(0.175) \given \text{data}) \geq 0.5$
\end{enumerate}

\subsection{Promising attempt 1}
This attempt keeps the two-stages design by introducing a virtual control arm, into which no patients are recruited. It uses N(-12,12) for treatment effect, and N(logit(0.075), 1/1000) as prior for control. It used delta (the changes in logit ORRs), similar to the Gsponer's example 2. This attempt also defined the success and futility criteria using the required responders from \citet{Roychoudhury2018} Table 4. Using above settings, the result gives 0 inconclusive values, which matches the Roychoudhury's results (Table 4 in the paper).\\
Success criterion: $\mathbb{P}(\delta > \logit(5/25) - \logit(0.075)) \geq 0.5$\\
Futility criterion: $\mathbb{P}(\delta < \logit(5/25) - \logit(0.075)) \geq 0.5$.
<<try2, echo=TRUE>>=
logit <- function(p) { log(p/(1-p)) }
expit <- function(x) { exp(x)/(1+exp(x)) }
p0 = 0.075
p1 = 0.175

prior_treatment_precision = 1/12^2

suc = logit(5/25) - logit(p0)
fut = logit(5/25) - logit(p0)
design <- gsbDesign(nr.stages=2,
 patients=cbind( c(0,0), c(0,25) ),
 sigma=2,
 criteria.success=rbind( c(NA,NA), c(suc,0.5) ),
 criteria.futility=rbind( c(NA,NA),c(fut,0.5) ),
 prior.control=c( logit(p0),1000),
 prior.treatment=c(-12,prior_treatment_precision))

simulation <- gsbSimulation(
  truth=list(logit(0.075), # true control
             c(logit(0.075)-logit(0.075), logit(0.125)-logit(0.075),
               logit(0.175)-logit(0.075), # true treatment
               logit(0.225)-logit(0.075), logit(0.275)-logit(0.075))),
 grid.type = "sliced",
 type.update = "per arm",
 nr.sim = 100000,
 warnings.sensitivity = 2000,
 seed = 1)
result <- gsb(design,simulation)
result
#tab(result, what="cumulative all")
#tab(result, what="sample size")
#plot(result, what="sample size", sliced=TRUE)
tab(result, what="cumulative all") # control and treatment columns in logit scale
plot(result, what="cumulative all", sliced=TRUE)
par(mfrow=c(1,2))
y1 = tab(result,what='futility')[,5]
y2 = tab(result,what='success')[,5]
y3 = 1 - y1 - y2
piE = expit(c(logit(0.075), logit(0.125),logit(0.175), # true treatment
               logit(0.225), logit(0.275)))
plot(piE,y1,type='l',ylim=c(0,1),ylab='probability') # futility
lines(piE,y2) # success
lines(piE,y3)
y3 = tab(result,what='sample size')[,5]
plot(piE,y3,type='l',ylab='expected sample size') 
@

This attempt turns the previous two-stage design to a one-stage design. 
<<try3, echo=TRUE>>=
logit <- function(p) { log(p/(1-p)) }
expit <- function(x) { exp(x)/(1+exp(x)) }
p0 = 0.075
p1 = 0.175
suc = logit(5/25) - logit(p0)
fut = logit(5/25) - logit(p0)
design <- gsbDesign(nr.stages=1,
 patients=c(0,25),
 sigma=2,
 criteria.success= c(suc,0.5),
 criteria.futility= c(fut,0.5),
 prior.control=c( logit(p0),1000),
 prior.treatment=c(-12,1/144))

simulation <- gsbSimulation(
  truth=list(logit(0.075), # true control
             c(logit(0.075)-logit(0.075), logit(0.125)-logit(0.075),
               logit(0.175)-logit(0.075), # true treatment
               logit(0.225)-logit(0.075), logit(0.275)-logit(0.075))),
 grid.type = "sliced",
 type.update = "per arm",
 nr.sim = 100000,
 warnings.sensitivity = 2000,
 seed = 1)
result <- gsb(design,simulation)
result
#tab(result, what="cumulative all")
#tab(result, what="sample size")
#plot(result, what="sample size", sliced=TRUE)
tab(result, what="cumulative all") # control and treatment columns in logit scale
plot(result, what="cumulative all", sliced=TRUE)
par(mfrow=c(1,2))
y1 = tab(result,what='futility')[,4]
y2 = tab(result,what='success')[,4]
y3 = 1 - y1 - y2
piE = expit(c(logit(0.075), logit(0.125),logit(0.175), # true treatment
               logit(0.225), logit(0.275)))
plot(piE,y1,type='l',ylim=c(0,1),ylab='probability') 
lines(piE,y2)
lines(piE,y3)
y3 = tab(result,what='sample size')[,4]
plot(piE,y3,type='l',ylab='expected sample size') 
@

This attempt look at Roychoundhury's second case in Table 4 (sample size = 36). The results is quite far away from Roychoudhury's values (especially in GO case).
<<try4, echo=TRUE>>=
logit <- function(p) { log(p/(1-p)) }
expit <- function(x) { exp(x)/(1+exp(x)) }
p0 = 0.075
p1 = 0.175
suc = logit(7/25) - logit(p0)
fut = logit(5/25) - logit(p0)
design <- gsbDesign(nr.stages=1,
 patients=c(0,36),
 sigma=2,
 criteria.success= c(suc,0.5),
 criteria.futility= c(fut,0.5),
 prior.control=c( logit(p0),1000),
 prior.treatment=c(-12,1/144))

simulation <- gsbSimulation(
  truth=list(logit(0.075), # true control
             c(logit(0.075)-logit(0.075), logit(0.125)-logit(0.075),
               logit(0.175)-logit(0.075), # true treatment
               logit(0.225)-logit(0.075), logit(0.275)-logit(0.075))),
 grid.type = "sliced",
 type.update = "per arm",
 nr.sim = 100000,
 warnings.sensitivity = 2000,
 seed = 1)
result <- gsb(design,simulation)
result
#tab(result, what="cumulative all")
#tab(result, what="sample size")
#plot(result, what="sample size", sliced=TRUE)
tab(result, what="cumulative all") # control and treatment columns in logit scale
plot(result, what="cumulative all", sliced=TRUE)
par(mfrow=c(1,2))
y1 = tab(result,what='futility')[,4]
y2 = tab(result,what='success')[,4]
y3 = 1 - y1 - y2
piE = expit(c(logit(0.075), logit(0.125),logit(0.175), # true treatment
               logit(0.225), logit(0.275)))
plot(piE,y1,type='l',ylim=c(0,1),ylab='probability') 
lines(piE,y2)
lines(piE,y3)
y3 = tab(result,what='sample size')[,4]
plot(piE,y3,type='l',ylab='expected sample size') 
@

\subsection{Not so successful attempts}
I made some attempts but met some problems too:\\
This attempt uses $\Nor(\logit(0.075),1/1000)$ as prior for control, and $\Nor(-12,12^2)$ as prior for treatment. It uses two futility criteria.
<<try0, echo=TRUE>>=
library(gsbDesign)
logit <- function(p) { log(p/(1-p)) }
expit <- function(x) { exp(x)/(1+exp(x)) }

# single-arm design
design <- gsbDesign(nr.stages=1,
 patients=c(0,25), 
 sigma=2,
 criteria.success=c(logit(0.075), 0.95, logit(0.175), 0.5),
 criteria.futility=c(logit(0.075), 0.05, logit(0.175), 0.5),
 prior.control=c(logit(0.075),1000), # very small variance in control prior
 prior.treatment=c(-12, 1/12^2)) # large variance in treatment prior

simulation <- gsbSimulation(
  truth=list(logit(0.075), # true control
             c(logit(0.075), logit(0.125),logit(0.175), # true treatment
               logit(0.225), logit(0.275))),
 grid.type = "table",
 type.update = "per arm", # only simulation method is implemented
 nr.sim = 100000,
 warnings.sensitivity = 2000,
 seed = 1)

simulation <- gsbSimulation(
  truth=list(logit(0.075), # true control
             c(logit(0.075), logit(0.125), logit(0.175), # true treatment
               logit(0.225), logit(0.275))),
 grid.type = "table",
 type.update = "per arm", # only simulation method is implemented
 nr.sim = 100000,
 warnings.sensitivity = 2000,
 seed = 1)
result <- gsb(design,simulation)
result
#tab(result, what="cumulative all")
#tab(result, what="sample size")
#plot(result, what="sample size", sliced = FALSE)
plot(result, what="cumulative all", sliced=FALSE)
par(mfrow=c(1,2))
y1 = tab(result,what='futility')[,4]
y2 = tab(result,what='success')[,4]
y3 = 1 - y1 - y2
piE = expit(c(logit(0.075), logit(0.125),logit(0.175), # true treatment
               logit(0.225), logit(0.275)))
plot(piE,y1,type='l',ylim=c(0,1),ylab='probability') 
lines(piE,y2)
lines(piE,y3)
y3 = tab(result,what='sample size')[,4]
plot(piE,y3,type='l',ylab='expected sample size') 
@

This attempt uses $\Nor$(-12, 1/1000) as prior for control, and $\Nor(-12,12^2)$ as prior for treatment. It only have one criteria for futility(corresponding to 1 stage in total), as the examples in Gsponer's paper. The result returns indeterminate situations.
<<try1, echo=TRUE>>=
library(gsbDesign)
logit <- function(p) { log(p/(1-p)) }
expit <- function(x) { exp(x)/(1+exp(x)) }

# single-arm design
design <- gsbDesign(nr.stages=1,
 patients=c(0,25), 
 sigma=2,
 criteria.success=c(logit(0.075), 0.95, logit(0.175), 0.5),
 criteria.futility=c(logit(0.075), 0.05),
 prior.control=c(-12,1000), # very small variance in control prior
 prior.treatment=c(-12, 1/12^2)) # large variance in treatment prior

simulation <- gsbSimulation(
  truth=list(-12, # true control
             c(logit(0.075), logit(0.125),logit(0.175), # true treatment
               logit(0.225), logit(0.275))),
 grid.type = "sliced",
 type.update = "per arm", # only simulation method is implemented
 nr.sim = 100000,
 warnings.sensitivity = 2000,
 seed = 1)

result <- gsb(design,simulation)
result
#tab(result, what="cumulative all")
#tab(result, what="sample size")
#plot(result, what="sample size", sliced = T)
plot(result, what="cumulative all", sliced=T)
par(mfrow=c(1,2))
y1 = tab(result,what='futility')[,4]
y2 = tab(result,what='success')[,4]
y3 = 1 - y1 - y2
piE = expit(c(logit(0.075), logit(0.125),logit(0.175), # true treatment
               logit(0.225), logit(0.275)))
plot(piE,y1,type='l',ylim=c(0,1),ylab='probability') 
lines(piE,y2)
lines(piE,y3)
y3 = tab(result,what='sample size')[,4]
plot(piE,y3,type='l',ylab='expected sample size') 
@

Below code finds the minimal sample size with the help of gsbDesign package.[Not finished]
<<minimal sample size search, eval=FALSE>>=

SS_Bayesian <- function(
  null.value,
  p.positive = 0.95,
  dec.value,
  mu = -12, 
  sigma = 2,
  n.max = 100
){
  NV=null.value
  DV=dec.value
  n <- 1:n.max

}

SS_DC_BayesBin <- function(
  null.value, # null.value
  p.positive  = 0.95, # required probability to be better than the null
  dec.value,  # decision value
  post.est = c("median","mean")[1], # median by default
  a=1, b=1,    # prior specification, a Beta(a,b) prior
  n.max = 1000 # numbers of iterations
){
  NV = null.value
  DV = dec.value
  n <- 1:n.max

  # required number of responders to satisfy DV criterion (posterior mean)
  if (post.est == "mean") { # if estimate posterior mean
    rr = rep(NA,max(n)) # a vector to store rr (required responders)
      for ( nn in n) { # n iterations
      r = 0:nn # different numbers of responders
      est1 = (a+r)/(a+b+nn) # posterior estimate based on different success numbers
      rr[nn] = r[est1>=DV][1] # minimum r s.t. DV criterion is fulfilled
      }
    print(cbind(rr,n)) # required responders for each sample size n 
    est = (a+rr)/(a+b+n) # posterior estimate of mean
  }

  # required number of responders to satisfy DV criterion (posterior median)
  if (post.est == "median") {
    rr = rep(NA,max(n))
    for ( nn in n) {
      r = 0:nn
      est1= qbeta(0.5,a+r,b+nn-r) # the median of beta posterior
      rr[nn] = r[est1>=DV][1] # minimum r s.t. posterior median >= DV
    }
    print(cbind(rr,n))
    est = qbeta(0.5,a+rr,b+n-rr)
  }

  p.obs = rr/n

  # find n such that posterior prob >= p.positive (= 0.95)
  # in other words, find n such that statistical significance is guaranteed.
  post.crit <- 1 - pbeta(NV,a+rr,b+n-rr) # posterior probability of being greater than NV.
  post.crit.index = post.crit >= p.positive # index=TRUE if the probability >= 0.95

  # find n such that for all m>=n, statistical criterion is fulfilled
  find.n =  cbind( n, rev(post.crit.index), 
                   cumsum(rev(post.crit.index))==n, 
                   # true if all m>=n fulfill stats criterion
                   rev(n))
  colnames(find.n) = c("index","okay","all.next.okay","n")

  n.min=NULL

  if (any(find.n[,3]==1) ) {
    n.okay = find.n[find.n[,3]==1,,drop=FALSE] # take all the rows with all.next.okay==1
    n.min = n.okay[nrow(n.okay),4] # take the n in the last row (find minimum n)
    n.min.15 = n.min+15
    est=round(est,3)
    ss.table = data.frame( n=n[1:n.min.15], 
                           # sample size from 1 to n.min+10
                           r=rr[1:n.min.15], 
                           # number of responders needed for DV criterion
                           p.obs=p.obs[1:n.min.15], 
                           est = est[1:n.min.15], 
                           # posterior estimate of median (by default)
                           post.p.positive=post.crit[1:n.min.15], 
                           # posterior probability of a positive effect
                           okay=post.crit.index[1:n.min.15]
                           # TRUE if statistical significance is satisfied
                           )}

  if (is.null(n.min))
    stop("No sample size satisfies required posterior 
         probability criterion pr(p>NV): try higher n.max")

plot( 1:(n.min+15), post.crit[1:(n.min+15)], type="p", xaxs="i", xlab="n",
        ylab="pr(p>NV)",xlim = c(0,n.min+20), ylim=c(0.5,1),
        lwd=1, pch = ifelse(post.crit.index,16,1),
        main="prob(p>NV) as a function of n", cex.lab=1.4, cex.main=1.4)
  abline(h=p.positive, lty=2)
  abline(v=n.min, lty=3)
  text(1.05*n.min, 0.6, paste("n.min=",n.min, sep=''),adj=0)
  text(30, 0.8, paste("a=",round(a,4)))

  return(list(ss.table=ss.table,n.min=n.min))}

@

%% References
\appendix
\section*{Appendix}

\bibliographystyle{apalike}
\bibliography{bibliography_1}

\newpage
\section*{Computational details}
<< "sessionInfo" >>=
cat(paste(Sys.time(), Sys.timezone(), "\n"))
sessionInfo()
@

\end{document}
